### 2023-02-22 paper/main copy.tex
init

% vim: ft=tex
\documentclass[a4,natbib=false]{article}
\usepackage[a4paper, total={8in, 10in}]{geometry}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{supertabular}
\usepackage[inline]{enumitem}
\usepackage{cleveref}

\newcommand{\niton}{\not\owns}

\newcommand{\ilyam}[1]{{\color{red} \texttt{Ilya:  #1}}}

\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}

\include{_build/_gen/included.tex}

\begin{document}

\begin{definition}[Context simplification]
  We simplify the context by applying it to itself.

  \ottfundefnsimpl{}
\end{definition}

\begin{lemma}
  Context is applicatively equivalent to its simplification.
  \[ [[ T* ]] \simeq [[ T ]] \]
\end{lemma}

\begin{definition}[Half-optimized unification algorithm]
  \label{def:half}
  I optimize the Ambroise's algorithm by simplifying the output context and
  removing the unused metavariables.

  \ottdefnun{}
\end{definition}

\begin{definition}[Fully-optimized unification algorithm]
  \label{def:full}
  I optimize the algorithm further by unrolling the UV\_F sub-inference into one step.

  \ottdefnaunTwo{}
\end{definition}

\begin{definition}[External terms]
  External terms are those whose metavariables do not have arguments bound by a
  constructor.

  \ottdefnaext{}
\end{definition}

\begin{definition}[External Contexts]
  External contexts are those whose assignments have external RHS.

  \ottdefnaextC{}
\end{definition}

\begin{lemma}
  \label{lemma:ext-app}
  if $ [[ t ext ]] $ and $ [[ T ext ]] $ then $ [[ [T]t ext ]] $.
\end{lemma}

\begin{lemma}
  Given a metavariable signature set  $\Sigma = [[ uc1[zs1] ]] \dots [[uc1[zsn] ]]$,
  an applied unification variable $[[ ua[xs] ]]$, an input context $[[ T1 ]]$,
  and a term $t \niton [[ua]]$ such that
  \begin{enumerate}
  \item $t$ and $[[ ua[xs] ]]$ are well-formed in $[[T1]]$
    % , i.e., $[[T1]]$ contains their unification variables and free variables.
  \item $[[T1]]$  is well-formed, external, and simplified ($ [[T1 ext]] $,  $[[T1*]] = [[T1]]$);
  \item $[[t]]$ is external ($[[t ext]]$);
  \item $[[ua : arn]] \in [[T1]]$
  \item $ [[T1]] $ is applied to $t$ ($[[ [T1] t ]] = [[t]]$);
  \item metavariable applications of $t$ are dictated by $\Sigma$
    (i.e. $UV(t) \subseteq \Sigma$), and $[[ ua[xs] ]] \in \Sigma$;
  \item for any $ [[ uci[zsi] ]] \in \Sigma $, the metavariable applications of
    $ [[ [T1](uci[zsi]) ]] $ are dictated by $\Sigma$
    (i.e. $UV([[ [T1]uci[zsi] ]]) \subseteq \Sigma$)
  \end{enumerate}
  if the half-optimized algorithm \ref{def:half} unifies $[[ ua[xs] ]]$ and
  $t$ in context $[[ T1 ]]$ with the resulting context $[[ T2 ]]$
  (i.e. $[[T1 ⊨ ua[xs] ≡ t : 0 ⫤ T2]]$),
  \textbf{then} the fully-optimized algorithm \ref{def:full}
  unifies them with certain resulting context $[[ T2' ]]$, equivalent to $[[ T2 ]]$
  (i.e. $[[T1 ⊫ ua[xs] ≡ t : 0 ⫥ T2']]$) and moreover,
  \begin{enumerate}
    \item $ [[ T2 ]] $ is well-formed, external and simplified;
    \item there exists $\Sigma' \supseteq \Sigma$ such that 
      $UV(\Sigma') \supseteq UV( [[ T2 ]] )$ and
      for any $ [[ uci[zsi] ]] \in \Sigma' $, the metavariable entries of 
      $ [[ [T2](uci[zsi]) ]] $ are dictated by $\Sigma'$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Induction on $t$.
  \begin{description}
  \item[$t = y$] In this case, $y = x_i$ and $UV(t) = \emptyset$.
    Then the output context of both algorithms is $[[(T1{ua:arn=xs.xi})*]]$.
    Let us prove that the output context satisfies the required properties.
    \begin{enumerate}
      \item This context is well-formed because $[[T1{ua:arn=xs.xi}]]$ is well-formed;
      \item This context is external because the input context $[[T1]]$ is
        external and the context update affects the part to the right from
        $[[ua:arn]]$ by applying $[[ua:arn=xs.xi]]$ (external!) to each
        instantiation (external!), which preserves the externality by
        \cref{lemma:ext-app}.
      \item This context is simplified by construction.
      \item Since no new metavariables are created, we keep the signature:
        $\Sigma' = \Sigma$. Since the input context $[[T1]]$ is simplified,
        the only instantiations that changed in the output context
        $[[(T1{ua:arn=xs.xi})*]]$ are $[[ua:arn]]$
        and those that contained $$[[ua]]$$. The substitution  
        $[[ua:arn=xs.xi]]$ only eliminates the entries of $[[ua]]$,
        and does not produces other entries. As such, the set of metavariable
        entries reduces, and each entry is still dictated by $\Sigma$ .
    \end{enumerate}

  \item[$t = [[ ub[ys] ]]$] 
    The output context of both algorithms is
    $[[T2]] = [[ (T1{uc:|zs|}{ua:arn=xs.uc[zs]}{ub:ark=ys.uc[zs]})* ]]$ where
    $[[zs = xs ∩ ys]]$ (a fresh metavariable $[[uc:|zs|]]$ is inserted just before 
    the earliest of two occurrences $[[ua]]$ and $[[ub]]$ in $[[T1]]$).
    Let us show that this context satisfies the required properties.
    \begin{enumerate}
      \item The context is well-formed because it was well-formed before
        simplification \ilyam{a lemma is required}.
      \item The context is external because so are the new instantiations, and
        context externality is preserved under simplification.
      \item The context is simplified by construction.
      \item We take $\Sigma' = \Sigma \cup \{ [[ uc[zs] ]] \}$.
        Let us take an entry from $\Sigma'$, which is affected by the context application.
        Then this entry is either 
        \begin{itemize}
        \item $[[ uai[xsi] ]]$ or $[[ ubi[ysi] ]]$, then the context application
          transforms it into $[[ uc[zs] ]] \in \Sigma'$;
        \item other $[[ udi[zsi] ]]$, then since $[[T1]]$ is simplified,
          $[[ [T2]udi[zsi] ]] = [[ [ ub:ark=ys.uc[zs] ][ ua:arn=xs.uc[zs]
          ]([T1]udi[zsi]) ]]$, which means that
          $UV([[ [T2]udi[zsi] ]]) \subseteq [[ [T1]udi[zsi] ]] \cup \{ [[ uc[zs] ]]\} \subseteq \Sigma'$. 
        \end{itemize}
    \end{enumerate}

  \item[$t = [[ f(ys1.t1,..,ysm.tm) ]]$] 
    blah

  \end{description}

\end{proof}

\begin{lemma}
  Given a signature of unification variables $\Sigma = [[ ua1[xs1] ]] \dots [[ ua1[xsn] ]]$,

  \textbf{if}
  \begin{enumerate}
    \item input context $ [[ T1 ]] $  is external and simplified ($ [[ T1 ext ]]
      $,  $ [[ T1 * ]] = [[ T1 ]] $);
    \item input terms $ [[ t1 ]] $ and $ [[ t2 ]] $ are external ($ [[ ti ext ]] $);
    \item $ [[ T1 ]] $ is applied to the input terms ($ [[ [T1] ti ]] = [[ ti ]] $);
    \item for any $ [[ uai ]] $, its entries in the terms are exactly as in
      $\Sigma$ ($UV(t_i) \subseteq \Sigma$);
    \item for any $ [[ uai[xsi] ]] \in \Sigma $, the metavariable entries of 
      $ [[ [T1](uai[xsi]) ]] $ are exactly as in $\Sigma$ ($UV([[ [T1]uai[xsi] ]]) \subseteq \Sigma$)
  \end{enumerate}

  \textbf{then} in result of the fully-optimized unification \ref{def:full},
    \begin{enumerate}
    \item the output context $ [[ T2 ]] $  is equivalent to those of the
      half-optimized unification \ref{def:half};
    \item $ [[ T2 ]] $ is external and simplified;
    \item output terms are external $ [[ [T2] ti ext ]] $;
    \item there exists $\Sigma' \supseteq \Sigma$ such that 
      $UV(\Sigma') \supseteq UV( [[ T2 ]] )$ and
      for any $ [[ uai[xsi] ]] \in \Sigma' $, the metavariable entries of 
      $ [[ [T2](uai[xsi]) ]] $ are exactly as in $\Sigma'$.
  \end{enumerate}

\end{lemma}

\begin{proof}

  % Induction on the inference tree corresponding to the step of the fully-optimized unification \ref{def:full}.
  % \begin{description}
  %   \item [$aV\_V$] In this case, the algorithms work identically, without
  %     changing the context. As such, the lemma conclusions hold almost immediately:
  %     \begin{enumerate}
  %       \item the rules of the two (deterministic!) algorithms are identical so are the output contexts;
  %       \item the output context is identical to the input context, which is
  %         external and simplified by the assumption;
  %       \item output terms are identical to the input terms, which are external;
  %       \item since the context has not changed, we chose $\Sigma' = \Sigma$.
  %     \end{enumerate}
  %   \item [$aF\_F$]
  %     Assume that $[[T0 ⊫ f(xs1.v1,..,xn.vn) ≡ f(xs1.w1,..,xsn.wn) : 0 ⫥ Tn]]$ and
  %     $[[ T0 ⊨ f(v1,..,vn) ≡ f(w1,..,wn) : 0 ⫤ Tn' ]]$. We want to show that
  %     $[[Tn]]$ is equivalent to $[[Tn']]$ and satisfies the required properties.
  %     By inversion, $[[T0, xs1 ⊫ v1 ≡ w1 : ark1 ⫥ T1, xs1]], \dots, [[Tn-1, xsn ⊫ [Tn-1]vn ≡ [Tn-1]wn : arkn ⫥ Tn, xsn]]$ and 
  %     $[[T0 ⊨ v1 ≡ w1 : ark1 ⫤ T1']], \dots, [[T'n-1 ⊨ [T'n-1]vn ≡ [T'n-1]wn : arkn ⫤ T'n]]$.
  %     Let us proceed by induction on $n$.

  %     For $n = 0$, the output context and the input context coincide, and the
  %     required property hold by the same reasons as in the previous case.

  %     If the  statement holds for $n-1$, then

  %     \begin{enumerate*}
  %     \item [(i)] $[[Tn-1]]$ is equivalent to $[[Tn-1']]$, and hence, they can be
  %       used interchangeably;
  %     \item [(ii)] $Tn-1$ is external and simplified;
  %     \item [(iii)] $T1$ are external
  %     \end{enumerate*}

  %      is equivalent to $[[T'n-1]]$ and

  %     Then by the IH,

  %     (and thus, inter)

  %     $[[T0 ⊨ v1 ≡ w1 : ark1 ⫤ T1]], \dots, [[Tn-1 ⊨ [Tn-1]vn ≡
  %     [Tn-1]wn : arkn ⫤ Tn]]$

  %     Induction on $n$.

  %     By inversion, $ [[T0, xs1 ⊫ v1 ≡ w1 : ark1 ⫥ T1, xs1 ]] $.
  %     First, let us show that the unification hypothesis is applicable to the
  %     context $ [[T0, xs1]] $, terms $ [[v1]]$ and $[[w1]]$, and the original $\Sigma$.
  %     \begin{enumerate}
  %       \item TODO 
  %     \end{enumerate}

  %     When unifying $ [[ xs1.v1 ]] $ and $ [[ xs1.w1 ]] $, the half-optimized
  %     unification ends up with $ [[ T0, xs1 ⊨ v1 ≡ w1 : ark1 ⫤ T1', xs1 ]] $.

  %     Then by the IH,
  %     \begin{enumerate*}
  %     \item [(i)] $[[T1]]$ is equivalent to $[[T1']]$, and hence, they can be
  %       used interchangeably;
  %     \item [(ii)] $T1$ is external and simplified;
  %     \item [(iii)] $T1$ is external and simplified;
  %     \end{enumerate*}

  %     $ [[ [T1]v1 ]] $ and $ [[ [T1]w1 ]] $ are external; there
  %     exists $\Sigma_1 \supseteq \Sigma$ stable under $ [[ T1' ]] $.

  %     Notice that the above is enough to apply the induction hypothesis to
  %     $ [[T1, xs2 ⊫ [T1]v2 ≡ [T1]w2 : ark2 ⫥ T2, xs2 ]] $ and $\Sigma_1$.
  %     Again, the half-optimized unification ends up with 
  %     $ [[ T1', xs2 ⊨ v2 ≡ w2 : ark2 ⫤ T2', xs1 ]] $ and since $[[T1]] ∼
  %     [[T2]]$, we apply the induction hypothesis and continue the process.

  %     % \begin{enumerate}

  %     %   \ref{def:half} will e
  %     %   By the induction hypothesis
  %     % \end{enumerate}
  % \end{description}
\end{proof}

\end{document}