### 2023-04-20 proofs/tex/au-lemmas.tex
Add equivalence soundness and completeness proofs

\begin{lemma}[Soundness of the anti-unification algorithm]
  \label{lemma:au-soundness}
\end{lemma}

\begin{lemma}[Completeness of the anti-unification algorithm]
  \label{lemma:au-completeness}
\end{lemma}

\begin{lemma}[Initiality of the anti-unification algorithm]
  \label{lemma:au-initiality}
\end{lemma}