### 2023-04-27 proofs/tex/anti-unification-lemmas.tex

\begin{lemma}[Soundness of Anti-Unification] \label{lemma:anti-unification-soundness}
    \hfill
    \begin{itemize}
        \item [$+$] 
        \item [$-$] 
    \end{itemize}
\end{lemma}

\begin{lemma}[Completeness of Anti-Unification] \label{lemma:anti-unification-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] 
        \item [$-$] 
    \end{itemize}
\end{lemma}