### 2023-05-02 proofs/tex/solution-merge-lemmas.tex

\begin{lemma} [Soundness of Solution Merge] \label{lemma:merge-soundness}
    Suppose that $[[us1 : Θ1]]$ and $[[us2 : Θ2]]$ 
    and $[[us1 & us2]]$ is defined.
    Then 
    \begin{enumerate}
        \item $[[us1 & us2 : Θ1 ∪ Θ2]]$,
        \item $[[Θ1 ⊢ us1 & us2 ⇒ us1]]$, and
        \item $[[Θ2 ⊢ us1 & us2 ⇒ us2]]$.
    \end{enumerate}
\end{lemma}

\begin{lemma} [Completeness and Initiality of Solution Merge] 
    \label{lemma:merge-completeness}
    Suppose that $[[us1 : Θ1]]$ and $[[us2 : Θ2]]$
    and there exists $[[us]]$
    such that $[[Θ1 ⊢ us1 ⇒ us]]$ and $[[ Θ2 ⊢ us2 ⇒ us]]$.
    Then $[[us1 & us2]]$ is defined and 
    $[[ Θ1 ∪ Θ2 ⊢ us1 & us2 ⇒ us]]$.
\end{lemma}

\begin{lemma} [Solution Weakening is Monotonous]
    \label{lemma:weakening-monotonicity}
    if $[[us1 : Θ]]$ and $[[us2 : Θ]]$ and 
    $[[us1 ⊆ us2]]$ then $[[Θ ⊢ us2 ⇒ us1]]$.
\end{lemma}

\begin{lemma} [Solution Weakening is transitive]
    \label{lemma:weakening-transitivity}
    Assuming $[[us1 : Θ]]$, $[[us2 : Θ]]$ and $[[us3 : Θ]]$,
    if $[[Θ ⊢ us1 ⇒ us2]]$ and $[[Θ ⊢ us2 ⇒ us3]]$ then $[[Θ ⊢ us1 ⇒ us3]]$.
\end{lemma}