### 2023-05-18 paper/tex/abstract.tex
Add soundness of solution merge.

My work focuses on bidirectional type inference. This involves combining the
checking and inference into one algorithm. During my first year, I utilized an
approach called polarization, also known as call-by-push-value. This method
separates the system into two subsystems: positive (values) and negative
(computations). This approach crystallizes the dichotomy and unveils the
underlying dualities not only between values and computations, but also between
polymorphism and existentialism, and algorithmically, unification and
anti-unification.