### 2023-05-23 proofs/tex/unification-lemmas.tex
Add soundness and completeness of unification Unification solution merge needs to be updated

\begin{lemma}[Soundness of Unification] \label{lemma:unification-soundness}
    \hfill
    \begin{itemize}
        \item [$+$] For normalized $[[uP]]$ and $[[iQ]]$ such that 
        $[[Γ ; Θ ⊢ uP]]$ and $[[Γ ⊢ iQ]]$,\\ 
        if $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us]]$ then $[[us : Θ | uv uP]]$, 
        $[[us]]$ is equivalence-only, and $[[ [us]uP = iQ ]]$.

        \item [$-$] For normalized $[[uN]]$ and $[[iM]]$ such that
        $[[Γ ; Θ ⊢ uN]]$ and $[[Γ ⊢ iM]]$,\\
        if $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$ then $[[us : Θ | uv uN]]$,
        $[[us]]$ is equivalence-only, and $[[ [us]uN = iM ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove by induction on the derivation of 
    $[[ Γ ; Θ ⊨ uN ≈u iM ⫤ us ]]$ and mutually $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us]]$.
    Let us consider the last rule forming this derivation. 
    \begin{caseof}
        \item \ruleref{\ottdruleUNVarLabel}, then $[[uN]] = [[α⁻]] = [[iM]]$.
        The resulting unification solution is empty: $[[us]] = [[·]]$.
        Since $[[uv uN]] = [[∅]]$, it satisfies $[[us : Θ | uv uP]]$.
        Vacuously, it is equivalence-only, and $[[ [us]α⁻ = α⁻ ]]$, that is $[[ [us]uN = iM ]]$.

        \item \ruleref{\ottdruleUShiftULabel}, then $[[uN]] = [[↑uP]]$ and $[[iM]] = [[↑iQ]]$.
        The algorithm makes a recursive call to $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us]]$ returning $[[us']]$.
        By induction hypothesis, $[[us' : Θ | uv uP]]$,
        and since $[[uv uN]] = [[uv ↑uP]] = [[uv uP]]$,  $[[us' : Θ | uv uN]]$.
        Finally, $[[ [us]uN ]] = [[ [us]↑uP ]] = [[ ↑[us]uP ]] = [[ ↑iQ ]] = [[ iM ]]$.

        \item \ruleref{\ottdruleUArrowLabel}, then $[[uN]] = [[uP → uN']]$ and $[[iM]] = [[iQ → iM']]$.
        The algorithm makes two recursive calls to $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us1]]$ and
        $[[Γ ; Θ ⊨ uN' ≈u iM' ⫤ us2]]$ returning $[[us1 & us2]]$ as the result.

        It is clear that $[[uP]]$, $[[uN']]$, $[[iQ]]$, and $[[iM']]$ are normalized,
        and that $[[Γ ; Θ ⊢ uP]]$, $[[Γ ; Θ ⊢ uN']]$, $[[Γ ⊢ iQ]]$, and $[[Γ ⊢ iM']]$.
        This way, the induction hypothesis is applicable to both recursive calls.
        By applying the induction hypothesis to $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us1]]$,
        we have:
        \begin{itemize}
            \item $[[us1 : Θ | uv uP]]$,
            \item $[[us1]]$ is equivalence-only, and
            \item $[[ [us1]uP = iQ ]]$.
        \end{itemize}
        By applying it to $[[Γ ; Θ ⊨ uN' ≈u iM' ⫤ us2]]$, we have:
        \begin{itemize}
            \item $[[us2 : Θ | uv uN']]$,
            \item $[[us2]]$ is equivalence-only, and
            \item $[[ [us2]uN' = iM' ]]$.
        \end{itemize}

        By the soundness of the unification solution merge (\cref{lemma:unif-sol-merge-soundness}), we have:
        \begin{itemize}
            \item $[[us1 & us2 : Θ | uv uP ∪ uv uN']]$, and hence, $[[us1 & us2 : Θ | uv uN]]$; 
            \item $[[us1 & us2]]$ is equivalence-only; 
            \item $[[us1 & us2]]$ is normalized;
            \item $[[us1 & us2 | uv uP]] = [[us1]]$ and hence, $[[ [us1 & us2]uP ]] = [[ [us1]uP ]] = [[iQ]]$;
            \item $[[us1 & us2 | uv uN']] = [[us2]]$ and hence, $[[ [us1 & us2]uN' ]] = [[ [us2]uN' ]] = [[iM']]$.
        \end{itemize}
        This way, $[[ [us1 & us2]uN ]] = [[ [us1 & us2]uP → [us1 & us2]uN' ]] = [[ iQ → iM' ]] = [[ iM ]]$.

        \item \ruleref{\ottdruleUForallLabel}, then $[[uN]] = [[∀pas.uN']]$ and $[[iM]] = [[∀pas.iM']]$.
        The algorithm makes a recursive call to $[[Γ,pas ; Θ ⊨ uN' ≈u iM' ⫤ us]]$
        returning $[[us]]$ as the result.

        The induction hypothesis is applicable: $[[Γ,pas ; Θ ⊢ uN']]$ and $[[Γ,pas ⊢ iM']]$ hold
        by inversion, and $[[uN']]$ and $[[iM']]$ are normalized, since $[[uN]]$ and $[[iM]]$ are.

        By the induction hypothesis, $[[us : Θ | uv uN']]$, then, since
        $[[uv uN]] = [[uv ∀pas.uN']] = [[uv uN']]$, we have $[[us : Θ | uv uN]]$.
        Finally, $[[ [us]uN ]] = [[ [us]∀pas.uN' ]] = [[ ∀pas.[us]uN' ]] = [[ ∀pas.iM' ]] = [[ iM ]]$.

        \item \ruleref{\ottdruleUNUVarLabel}, then $[[uN]] = [[α̂⁻]]$, $[[â⁻[Δ] ∊ Θ]]$, and $[[Δ ⊢ iM]]$.
        As the result, the algorithm returns $[[us]] = [[ (â⁻ :≈ iM) ]]$.

        It is clear that $[[(â⁻ :≈ iM) : (α̂⁻[Δ])]]$, since $[[Δ ⊢ iM]]$.
        Notice that since $[[Θ|uv uN]] = [[Θ|{α̂⁻}]] = [[(α̂⁻[Δ])]]$, we can rewrite 
        $[[(â⁻ :≈ iM) : (α̂⁻[Δ])]]$ as $[[us : Θ | uv uN]]$, as required.

        It is also clear that $[[us]] = [[ (â⁻ :≈ iM) ]]$ is equivalence-only, 
        and that $[[ [us]uN ]] = [[ [â⁻ :≈ iM]α̂⁻ ]] = [[ iM ]]$, as required.

        \item The positive cases are proved symmetrically.

    \end{caseof}

\end{proof}

\begin{lemma}[Completeness of Unification] \label{lemma:unification-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] For normalized $[[uP]]$ and $[[iQ]]$ such that
        $[[Γ ; Θ ⊢ uP]]$ and $[[Γ ⊢ iQ]]$,\\ 
        assume there exists an equivalence-only unification solution
        $[[us : Θ | uv uP]]$ such that $[[ [us]uP = iQ ]]$.
        Then $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.

        \item [$-$] For normalized $[[uN]]$ and $[[iM]]$ such that
        $[[Γ ; Θ ⊢ uN]]$ and $[[Γ ⊢ iM]]$,\\
        assume there exists an equivalence-only unification solution
        $[[us : Θ | uv uN]]$        
        such that $[[ [us]uN = iM ]]$.
        Then $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.
   \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on the structure of $[[uP]]$ and mutually, $[[uN]]$.
    \begin{caseof}
        \item $[[uN]] = [[α̂⁻]]$\\
            Since $[[uv uN]]$ = $[[α̂⁻]]$, 
            $[[us : Θ | {α̂⁻}]]$ means that $[[ α̂⁻[Δ] ]] \in [[Θ]]$, 
            and since $[[us]]$ is equivalence only, $[[us]] = [[ (α̂⁻ :≈ iM') ]]$
            for some $[[Δ ⊢ iM']]$.
            $[[ [us]uN = iM ]]$ means $[[iM' = iM]]$.
            This way, \ruleref{\ottdruleUNUVarLabel} is applicable to infer 
            $[[Γ ; Θ ⊨ â⁻ ≈u iM ⫤ (â⁻ :≈ iM)]]$.
        \item $[[uN]] = [[α⁻]]$\\
            Then $[[us : Θ | {α⁻}]]$ means that $[[ us ]] = [[·]]$, 
            and $[[ [us]α⁻ = iM ]]$ means $[[iM = α⁻]]$.
            This way, \ruleref{\ottdruleUNVarLabel} infers 
            $[[Γ; Θ ⊨ a⁻ ≈u a⁻ ⫤ ·]]$, which is rewritten as $[[Γ; Θ ⊨ uN ≈u iM ⫤ us]]$.

        \item $[[uN]] = [[↑uP]]$\\
            Notice that $[[ [us]uN = iM ]]$ means $[[ ↑[us]uP = iM ]]$, i.e.
            $[[iM]] = [[↑iQ]]$ for some $[[iQ]]$ and $[[ [us]uP = iQ ]]$.

            Let us show that the induction hypothesis is applicable to $[[ [us]uP = iQ ]]$.
            Notice that $[[uP]]$ is normalized, since $[[uN]] = [[↑uP]]$ is normalized,
            $[[Γ ; Θ ⊢ uP]]$ follows from inversion of $[[Γ ; Θ ⊢ ↑uP]]$, 
            and $[[Γ ⊢ iQ]]$ follows from inversion of $[[Γ ⊢ ↑iQ]]$,
            $[[us]]$ is equivalence-only by assumption, and
            $[[us : Θ | uv uP]]$, since $[[uv uP]] = [[uv ↑uP]] = [[uv uN]]$.

            Then by the induction hypothesis, $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us]]$.
            Hence, \ruleref{\ottdruleUShiftULabel} is applicable to infer
            $[[Γ ; Θ ⊨ ↑uP ≈u ↑iQ ⫤ us]]$, that is $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.

        \item $[[uN]] = [[uP → uN']]$\\
            Notice that $[[ [us]uN = iM ]]$ means $[[ [us]uP → [us]uN' = iM ]]$, i.e.
            $[[iM]] = [[iQ → iM']]$ for some $[[iQ]]$ and $[[iM']]$, 
            such that $[[ [us]uP = iQ ]]$ and $[[ [us]uN' = iM' ]]$.
            Then $[[ [us | uv uP]uP = iQ ]]$ and $[[ [us | uv uN']uN' = iM' ]]$.

            Let us show that the induction hypothesis is applicable to $[[ [us | uv uP]uP = iQ ]]$ 
            and $[[ [us | uv uN']uN' = iM' ]]$.
            Notice that $[[uP]]$ and $[[uN']]$ are normalized, since $[[uN]] = [[uP → uN']]$ is normalized,
            $[[Γ ; Θ ⊢ uP]]$ and $[[Γ ; Θ ⊢ uN']]$ follow from inversion of $[[Γ ; Θ ⊢ uP → uN']]$,
            $[[Γ ⊢ iQ]]$ and $[[Γ ⊢ iM']]$ follow from inversion of $[[Γ ⊢ iQ → iM']]$,
            $[[us|uv uP]]$ and $[[us|uv uN']]$ are equivalence-only since $[[us]]$ is equivalence-only,
            $[[us|uv uP : Θ | uv uP]]$ and $[[us|uv uN' : Θ | uv uN']]$, 
            since $[[us : Θ | uv uP ∪ uv uN']]$ (\cref{todo}).

            Then by the induction hypothesis, $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us|uv uP]]$ and
            $[[Γ ; Θ ⊨ uN' ≈u iM' ⫤ us|uv uN']]$.
            To apply \ruleref{\ottdruleUArrowLabel} and infer the required
            $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$, we need to show that
            $[[us|uv uP & us|uv uN']] = [[us]]$. 
            By the completeness of the unification solution merge 
            (\cref{lemma:unif-sol-merge-completeness}), 
            $[[us|uv uP & us|uv uN']] = [[us|uv uP ∪ uv uN']]$,
            and $[[us|uv uP ∪ uv uN']] = [[us|uv uN]] = [[us]]$, since $[[us : Θ | uv uN]]$.

        \item $[[uN]] = [[∀pas.uN']]$\\
            Notice that $[[ [us]uN = iM ]]$ means $[[ ∀pas.[us]uN' = iM ]]$, i.e.
            $[[iM]] = [[∀pas.iM']]$ for some $[[iM']]$ such that $[[ [us]uN' = iM' ]]$.

            Let us show that the induction hypothesis is applicable to $[[ [us]uN' = iM' ]]$.
            Notice that $[[uN']]$ is normalized, since $[[uN]] = [[∀pas.uN']]$ is normalized,
            $[[Γ,pas ; Θ ⊢ uN']]$ follows from inversion of $[[Γ ; Θ ⊢ ∀pas.uN']]$,
            $[[Γ,pas ⊢ iM']]$ follows from inversion of $[[Γ ⊢ ∀pas.iM']]$,
            $[[us]]$ is equivalence-only by assumption, and
            $[[us : Θ | uv uN']]$ since $[[uv uN']] = [[uv ∀pas.uN']] = [[uv uN]]$.

            By the induction hypothesis, $[[Γ,pas ; Θ ⊨ uN' ≈u iM' ⫤ us]]$.
            Hence, \ruleref{\ottdruleUForallLabel} is applicable to infer
            $[[Γ ; Θ ⊨ ∀pas.uN' ≈u ∀pas.iM' ⫤ us]]$, that is $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.

        \item The positive cases are proved symmetrically.
    \end{caseof}
\end{proof}