### 2023-07-10 proofs/tex/decl-typing-lemmas.tex
wip

\begin{lemma} \label{lemma:app-inf-equ-stable}
    If $[[Γ; Φ ⊢ iN1 ● args ⇒> iM]]$ and $[[Γ ⊢ iN1 ≈ iN2]]$ 
    then $[[Γ; Φ ⊢ iN2 ● args ⇒> iM]]$.
\end{lemma}
\begin{proof}
    By \cref{lemma:equiv-completeness}, 
    $[[Γ ⊢ iN1 ≈ iN2]]$ implies $[[iN1 ≈ iN2]]$.
    Let us prove the required judgement by induction on $[[iN1 ≈ iN2]]$.
    Let us consider the last rule used in the derivation.
    \begin{caseof}
        \item \ruleref{\ottdruleEOneNVarLabel}.
            It means that $[[iN1]]$ is $[[α⁻]]$ and $[[iN2]]$ is $[[α⁻]]$.
            Then the required property coincides with the assumption. 
        \item \ruleref{\ottdruleEOneShiftULabel}. 
            It means that $[[iN1]]$ is $[[↑iP1]]$ and $[[iN2]]$ is $[[↑iP2]]$.
            where $[[iP1 ≈ iP2]]$.

            Then the only rule applicable to infer $[[Γ; Φ ⊢ ↑iP1 ● args ⇒> iM]]$
            is \ruleref{\ottdruleDTEmptyAppLabel},
            meaning that $[[args = ·]]$ and $[[Γ ⊢ ↑iP1 ≈ iM]]$.
            Then by transitivity of equivalence \cref{corollary:equivalence-transitivity},
            $[[Γ ⊢ ↑iP2 ≈ iM]]$, and then \ruleref{\ottdruleDTEmptyAppLabel} is applicable to infer
            $[[Γ; Φ ⊢ ↑iP2 ● · ⇒> iM]]$.

        \item \ruleref{\ottdruleEOneArrowLabel}.
            Then we are proving that  
            $[[Γ; Φ ⊢ (iQ1 → iN1) ● v, args ⇒> iM]]$ and $[[iQ1 → iN1 ≈ iQ2 → iN2]]$
            imply $[[Γ; Φ ⊢ (iQ2 → iN2) ● v, args ⇒> iM]]$.

            By inversion, $[[(iQ1 → iN1) ≈ (iQ2 → iN2)]]$
            means $[[iQ1 ≈ iQ2]]$ and $[[iN1 ≈ iN2]]$.

            By inversion of $[[Γ; Φ ⊢ (iQ1 → iN1) ● v, args ⇒> iM]]$:
            \begin{enumerate}
                \item $[[Γ ; Φ ⊢ v : iP]]$
                \item $[[Γ ⊢ iQ1 ≥ iP]]$,
                    and then by transitivity \cref{corollary:subtyping-transitivity},
                    $[[Γ ⊢ iQ2 ≥ iP]]$;
                \item $[[Γ ; Φ ⊢ iN1 ● args ⇒> iM]]$, 
                    and then by induction hypothesis, $[[Γ ; Φ ⊢ iN2 ● args ⇒> iM]]$.
            \end{enumerate}

            Since we have $[[Γ ; Φ ⊢ v : iP]]$, $[[Γ ⊢ iQ2 ≥ iP]]$ and 
            $[[Γ ; Φ ⊢ iN2 ● args ⇒> iM]]$, we can apply \ruleref{\ottdruleDTArrowAppLabel}
            to infer $[[Γ; Φ ⊢ (iQ2 → iN2) ● v, args ⇒> iM]]$.

        \item \ruleref{\ottdruleEOneForallLabel}
            Then we are proving that 
            $[[Γ ; Φ ⊢ ∀pas1.iN1' ● args ⇒> iM]]$ and $[[∀pas1.iN1' ≈ ∀pas2.iN2']]$
            imply $[[Γ ; Φ ⊢ ∀pas2.iN2' ● args ⇒> iM]]$.

            By inversion of $[[∀pas1.iN1' ≈ ∀pas2.iN2']]$:
            \begin{enumerate}
                \item $[[{pas2} ∩ fv iN1 = ∅]]$,
                \item there exists a bijection 
                    $[[mu : ({pas2} ∩ fv iN2') ↔ ({pas1} ∩ fv iN1')]]$
                    such that $[[iN1' ≈ [mu] iN2']]$.
            \end{enumerate}

            By inversion of $[[Γ ; Φ ⊢ ∀pas1.iN1' ● args ⇒> iM]]$:
            \begin{enumerate}
                \item $[[Γ ⊢ σ : pas1]]$        
                \item $[[Γ ; Φ ⊢ [σ]iN1' ● args ⇒> iM]]$
                \item $[[args ≠ ·]]$
            \end{enumerate}

            Let us construct $[[Γ ⊢ σ0 : pas2]]$ in the following way:
            $$
            \begin{cases}
                [[ [σ0]α⁺ =  [σ][mu]α⁺ ]] & \text{if } [[α⁺]] \in [[ {pas2} ∩ fv iN2' ]] \\
                [[ [σ0]α⁺ =  ∃β⁻.↓β⁻ ]] & \text{otherwise (the type does not matter here)} \\
            \end{cases}
            $$

            Then to infer $[[Γ ; Φ ⊢ iN2 ● args ⇒> iM]]$, we 
            apply \ruleref{\ottdruleDTArrowAppLabel} with $[[σ0]]$. 
            Let us show the required premises:
            \begin{enumerate}
                \item $[[Γ ⊢ σ0 : pas2]]$ by construction;
                \item $[[args ≠ ·]]$ as noted above;
                \item To show $[[Γ ; Φ ⊢ [σ0]iN2' ● args ⇒> iM]]$,
                Notice that $[[ [σ0]iN2' = [σ][mu]iN2' ]]$   
                and since $[[ [mu]iN2' ≈ iN1' ]]$, $[[ [σ][mu]iN2' ≈ [σ]iN1' ]]$.
                This way, by \cref{lemma:equiv-soundness}, $[[Γ ⊢ [σ]iN1' ≈ [σ0]iN2']]$.
                Then the required judgement holds by the induction hypothesis
                applied to $[[Γ ; Φ ⊢ [σ]iN1' ● args ⇒> iM]]$.
            \end{enumerate}
    \end{caseof}
\end{proof}

\newcommand{\pureSize}[1]{\ensuremath{\mathsf{pure\_size}(#1)}}
\newcommand{\eqNodes}[1]{\ensuremath{\mathsf{eq\_nodes}(#1)}}
\newcommand{\size}[1]{\ensuremath{\mathsf{size}(#1)}}

\begin{definition}
    For a tree $T$ inferring
    a declarative typing judgement 
    (i.e. the one of the shape
    $[[Γ ; Φ ⊢ v : iP]]$, 
    $[[Γ ; Φ ⊢ c : iN]]$, or 
    $[[Γ ; Φ ⊢ iN ● args ⇒> iM]]$),
    let us define a metrics $\size{T}$
    as the number of nodes in $T$ that 
    have one of these shapes
    (i.e., $[[Γ ; Φ ⊢ v : iP]]$, 
    $[[Γ ; Φ ⊢ c : iN]]$, or 
    $[[Γ ; Φ ⊢ iN ● args ⇒> iM]]$).
\end{definition}

\begin{definition}
    For a tree $T$ inferring
    a declarative typing judgement,
    let us a function $\eqNodes{T}$
    as the number of nodes in $T$ labeled with \ruleref{\ottdruleDTPEquivLabel} or 
    \ruleref{\ottdruleDTNEquivLabel}.
\end{definition}

\begin{definition}
    For a tree $T$ inferring
    a declarative typing judgement 
    let us define a metrics $\pureSize{T}$
    as the number of nodes in $T$ except 
    the ones labeled with \ruleref{\ottdruleDTPEquivLabel} or 
    \ruleref{\ottdruleDTNEquivLabel}.
    In other words, $\pureSize{T} = \size{T} - \eqNodes{T}$. 
\end{definition}

\begin{lemma}[Declarative typing is preserved under context equivalence]
    Assuming $[[Γ ⊢ Φ1]]$, $[[Γ ⊢ Φ2]]$, and $[[Γ ⊢ Φ1 ≈ Φ2]]$:
    \begin{itemize}
        \item [$+$] 
            for any tree $T_1$ inferring $[[Γ ; Φ1 ⊢ v : iP]]$, 
            there exists a tree $T_2$ inferring $[[Γ ; Φ2 ⊢ v : iP]]$ such that 
            $\pureSize{T_2} = \pureSize{T_1}$.
        \item [$-$] 
            for any tree $T_1$ inferring $[[Γ ; Φ1 ⊢ c : iN]]$, 
            there exists a tree $T_2$ inferring $[[Γ ; Φ2 ⊢ c : iN]]$ such that 
            $\pureSize{T_2} = \pureSize{T_1}$.
        \item [$\bullet$] 
            for any tree $T_1$ inferring $[[Γ ; Φ1 ⊢ iN ● args ⇒> iM]]$, 
            there exists a tree $T_2$ inferring $[[Γ ; Φ2 ⊢ iN ● args ⇒> iM]]$ such that 
            $\pureSize{T_2} = \pureSize{T_1}$.
    \end{itemize}
\end{lemma}
\begin{proof}
    Let us prove it by induction on a pair $(\pureSize{T_1}, \eqNodes{T_1})$.
    Let us consider the last rule applied in $T_1$ (i.e., its root node).
    \begin{caseof}
        \item \ruleref{\ottdruleDTVarLabel}\\
            Then we are proving 
            that $[[Γ ; Φ1 ⊢ x : iP]]$ implies $[[Γ ; Φ2 ⊢ x : iP]]$.
            By inversion, $[[x : iP ∊ Φ1]]$, and 
            since $[[Γ ⊢ Φ1 ≈ Φ2]]$, $[[x : iP' ∊ Φ2]]$ for some $[[iP']]$ 
            such that $[[Γ ⊢ iP ≈ iP']]$.
            Then we infer $[[Γ ; Φ2 ⊢ x : iP']]$ by \ruleref{\ottdruleDTVarLabel},
            and next, $[[Γ ; Φ2 ⊢ x : iP]]$ by \ruleref{\ottdruleDTPEquivLabel}.

            Notice that 
            to infer $T_1$, only \ruleref{\ottdruleDTVarLabel} was applied, 
            to infer $T_2$, only \ruleref{\ottdruleDTVarLabel} and
            \ruleref{\ottdruleDTPEquivLabel} were applied.
            This way, $\pureSize{T_1} = \pureSize{T_2} = 1$.

        \item For \ruleref{\ottdruleDTThunkLabel},
              \ruleref{\ottdruleDTPAnnotLabel}, 
              \ruleref{\ottdruleDTTLamLabel},
              \ruleref{\ottdruleDTReturnLabel},
              \ruleref{\ottdruleDTNAnnotLabel}, and
              the proof is analogous. We
              apply the induction hypothesis to the premise of the rule
              to substitute $[[Φ1]]$ for $[[Φ2]]$ in it. 
              The induction is applicable as the pure size of the 
              premises is less than the pure size of the conclusion by one.
              And after that, we apply the same rule to infer the required judgement.

              By induction hypothesis, 
              for each of the premises (i.e. the children $T_1'$ of $T_1$),
              the tree $T_2'$ inferring the corresponding judgement
              (with $[[Φ1]]$ substituted for $[[Φ2]]$) has the same pure size,
              i.e., $\pureSize{T_1'} = \pureSize{T_2'}$.
              For each of these rules, the pure size of $T_1$ 
              and the pure size $T_2$ is a sum of the pure sizes of their children
              plus one. This way, 
              $ 
                \pureSize{T_1} =
                1 + \sum_{T'_1}{\pureSize{T'_1}} =
                1 + \sum_{T'_2}{\pureSize{T'_2}} =
                \pureSize{T_2} 
              $.

        \item \ruleref{\ottdruleDTPEquivLabel} and \ruleref{\ottdruleDTNEquivLabel}
            In these cases, the induction hypothesis is also applicable to the premise:
            although the pure size of the premise is the same as the pure size of the conclusion,
            but the number of equivalence nodes is less by one.
            After this note, the proof is analogous to the previous case.

        \item \ruleref{\ottdruleDTtLamLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ λx:iP.c : iP → iN]]$ implies $[[Γ ; Φ2 ⊢ λx:iP.c : iP → iN]]$.
            Analogously to the previous case, 
            we apply the induction hypothesis to the
            equivalent contexts $[[Γ ⊢ Φ1, x:iP ≈ Φ2, x:iP]]$
            and the premise $[[Γ ; Φ1, x:iP ⊢ c : iN]]$
            inferred by $T'_1$ to obtain
            $T'_2$ inferring $[[Γ ; Φ2, x:iP ⊢ c : iN]]$.
            Then we construct $T_2$ inferring $[[Γ ; Φ2 ⊢ λx:iP.c : iP → iN]]$ by \ruleref{\ottdruleDTtLamLabel}
            and notice that $\pureSize{T_1} = 1 + \pureSize{T'_1} = 1 + \pureSize{T'_2} = \pureSize{T_2}$.

        \item \ruleref{\ottdruleDTVarLetLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ let x = v; c : iN]]$ implies $[[Γ ; Φ2 ⊢ let x = v; c : iN]]$.
            First, we apply the induction hypothesis to 
            $[[Γ; Φ1 ⊢ v : iP]]$ to obtain $[[Γ; Φ2 ⊢ v : iP]]$ 
            of the same pure size.

            Then we apply the induction hypothesis to
            the equivalent contexts $[[Γ ⊢ Φ1, x:iP ≈ Φ2, x:iP]]$
            and the premise $[[Γ ; Φ1, x:iP ⊢ c : iN]]$ to obtain
            $[[Γ ; Φ2, x:iP ⊢ c : iN]]$.
            Then we infer $[[Γ ; Φ2 ⊢ let x = v; c : iN]]$ by \ruleref{\ottdruleDTVarLetLabel}.

            The pure size of the resulting tree $T_2$ and of $T_1$ is
            one plus the sum of sizes of the premise trees.
            The premises of $T_1$ and of $T_2$ have the same pure size by 
            the induction hypothesis. This way, $\pureSize{T_1} = \pureSize{T_2}$.

        \item \ruleref{\ottdruleDTAppLetLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ let x = v(args); c : iN]]$ implies 
            $[[Γ ; Φ2 ⊢ let x = v(args); c : iN]]$.

            We apply the induction hypothesis to each of the premises.
            to rewrite:
            \begin{itemize}
                \item $[[Γ ; Φ1 ⊢ v : ↓iM]]$ into $[[Γ ; Φ2 ⊢ v : ↓iM]]$,
                \item $[[Γ ; Φ1 ⊢ iM ● args ⇒> ↑iQ]]$ into $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ]]$.
                \item $[[Γ ; Φ1, x:iQ ⊢ c : iN]]$ into $[[Γ ; Φ2, x:iQ ⊢ c : iN]]$
                (notice that $[[Γ ⊢ Φ1, x:iQ ≈ Φ2, x:iQ]]$).
            \end{itemize}

            It is left to show the uniqueness of $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ]]$.

            Then we infer $[[Γ ; Φ2 ⊢ let x = v(args); c : iN]]$ by \ruleref{\ottdruleDTAppLetLabel}.

        \item \ruleref{\ottdruleDTAppLetAnnLabel}
            Then we are proving that
            $[[Γ ; Φ1 ⊢ let x:iP = v(args); c : iN]]$ implies
            $[[Γ ; Φ2 ⊢ let x:iP = v(args); c : iN]]$.

            As in the previous case, we apply the induction hypothesis to each of the premises
            and rewrite:
            \begin{itemize}
                \item $[[Γ ; Φ1 ⊢ v : ↓iM]]$ into $[[Γ ; Φ2 ⊢ v : ↓iM]]$,
                \item $[[Γ ; Φ1 ⊢ iM ● args ⇒> ↑iQ]]$ into $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ]]$, 
                    and
                \item $[[Γ ; Φ1, x:iP ⊢ c : iN]]$ into $[[Γ ; Φ2, x:iP ⊢ c : iN]]$
                (notice that $[[Γ ⊢ Φ1, x:iP ≈ Φ2, x:iP]]$).
            \end{itemize}

            Notice that $[[Γ ⊢ iP]]$ and $[[Γ ⊢ ↑iQ ≤ ↑iP]]$ 
            do not depend on the variable context, and hold by assumption.
            Then we infer $[[Γ ; Φ2 ⊢ let x:iP = v(args); c : iN]]$ by \ruleref{\ottdruleDTAppLetAnnLabel}.

        \item \ruleref{\ottdruleDTUnpackLabel}, \ruleref{\ottdruleDTNAnnotLabel}, and \ruleref{\ottdruleDTNEquivLabel}
            are proved in the same way.

        \item \ruleref{\ottdruleDTEmptyAppLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ iN ● · ⇒> iN']]$ (inferred by \ruleref{\ottdruleDTEmptyAppLabel})
            implies $[[Γ ; Φ2 ⊢ iN ● · ⇒> iN']]$,
            and if there is only one $[[iN']]$ (up-to-equivalence in $[[Γ]]$) 
            that satisfies $[[Γ ; Φ1 ⊢ iN ● · ⇒> iN']]$,
            then there is only one $[[iN']]$ satisfying $[[Γ ; Φ2 ⊢ iN ● · ⇒> iN']]$.

            To infer $[[Γ ; Φ2 ⊢ iN ● · ⇒> iN']]$, 
            we apply \ruleref{\ottdruleDTEmptyAppLabel}, noting that 
            $[[Γ ⊢ iN ≈ iN']]$ holds by assumption.

            To show the uniqueness $[[Γ ; Φ2 ⊢ iN ● · ⇒> iN' uniq]]$, 
            let us assume $[[Γ ; Φ2 ⊢ iN ● · ⇒> iN'']]$.
            Then the only rule applicable to infer it is \ruleref{\ottdruleDTEmptyAppLabel},
            which, by inversion, means $[[Γ ⊢ iN ≈ iN'']]$, and by transitivity, 
            $[[Γ ⊢ iN' ≈ iN'']]$.

        \item \ruleref{\ottdruleDTArrowAppLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ iQ → iN ● v, args ⇒> iM]]$ (inferred by \ruleref{\ottdruleDTArrowAppLabel})
            implies $[[Γ ; Φ2 ⊢ iQ → iN ● v, args ⇒> iM]]$.
            And uniqueness of the $[[iM]]$ in the first case implies uniqueness in the second case.

            By induction, we rewrite $[[Γ ; Φ1 ⊢ v : iP]]$ into $[[Γ ; Φ2 ⊢ v : iP]]$, 
            and $[[Γ ; Φ1 ⊢ iN ● args ⇒> iM]]$ into $[[Γ ; Φ2 ⊢ iN ● args ⇒> iM]]$.
            Then we infer $[[Γ ; Φ2 ⊢ iQ → iN ● v, args ⇒> iM]]$ by \ruleref{\ottdruleDTArrowAppLabel}.

            Now, let us show the uniqueness.
            The only rule that can infer $[[Γ ; Φ1 ⊢ iQ → iN ● v, args ⇒> iM]]$
            is \ruleref{\ottdruleDTArrowAppLabel}.
            Then by inversion, 
            uniqueness of $[[Γ ; Φ1 ⊢ iQ → iN ● v, args ⇒> iM]]$ implies
            uniqueness of $[[Γ ; Φ1 ⊢ iN ● args ⇒> iM]]$. By 
            the induction hypothesis, it implies the uniqueness of 
            $[[Γ ; Φ2 ⊢ iN ● args ⇒> iM]]$.

            Suppose that 
            $[[Γ ; Φ2 ⊢ iQ → iN ● v, args ⇒> iM']]$.
            By inversion, $[[Γ ; Φ2 ⊢ iN ● args ⇒> iM']]$, 
            which by uniqueness of $[[Γ ; Φ2 ⊢ iN ● args ⇒> iM]]$ implies
            $[[Γ ⊢ iM ≈ iM']]$.

        \item \ruleref{\ottdruleDTForallAppLabel}
    \end{caseof}

\end{proof}