### 2023-09-05 proofs/tex/equiv-alg-lemmas.tex
Rearrange proofs

\begin{lemma}[Algorithmization of declarative equivalence]
  \label{lemma:decl-equiv-algorithmization}
  Declarative equivalence is the equality of normal forms. 
  \begin{itemize}
    \item[$+$] $[[iP ≈ iQ]] \iff [[nf(iP) = nf(iQ)]]$,
    \item[$-$] $[[iN ≈ iM]] \iff [[nf(iN) = nf(iM)]]$.
  \end{itemize}
\end{lemma}
\begin{proof} \hfill
  \begin{itemize}
    \item[$+$] Let us prove both directions separately.
    \begin{itemize}
      \item[$\Rightarrow$] 
        exactly by \cref{lemma:normalization-completeness},
      \item[$\Leftarrow$] 
        from \cref{lemma:normalization-soundness}, we know
        $[[iP ≈ nf(iP)]] = [[nf(iQ) ≈ iQ]]$, then by transitivity (\cref{lemma:decl-equiv-transitivity}),
        $[[iP ≈ iQ]]$.
    \end{itemize}
    \item[$-$] The proof is exactly the same.
  \end{itemize}
\end{proof}

\begin{lemma}[Algorithmic type well-formedness is invariant under equivalence]
  \label{lemma:wf-equiv-algo}
  Mutual subtyping implies declarative equivalence.
  \begin{itemize}
  \item[$+$] if $[[uP ≈ uQ]]$ then $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ uQ]]$,
  \item[$-$] if $[[uN ≈ uM]]$ then $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ uM]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof coincides with the proof of \cref{lemma:wf-equiv},
  and adds two cases for equating two positive or two negative algorithmic variables,
  which must be equal by inversion, and thus, 
  $[[Γ; Ξ ⊢ α̂±]] \iff [[Γ; Ξ  ⊢ α̂±]]$ holds trivially.
\end{proof}

\begin{corollary}[Normalization preserves well-formedness of algorithmic types]
  \label{corollary:wf-nf-algo}
  \begin{itemize}
    \item[$+$] $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ nf(uP)]]$,
    \item[$-$] $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ nf(uN)]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:wf-equiv-algo,lemma:normalization-soundness-alg}.
\end{proof}

\begin{corollary}[Normalization preserves the signature of the algorithmic substitution]
  \label{corollary:norm-subst-sig-algo}
  \hfill \\
  $[[Θ ⊢ uσ : Ξ]] \iff [[Θ ⊢ nf(uσ) : Ξ]]$,
  $[[Γ ⊢ uσ : Ξ]] \iff [[Γ ⊢ nf(uσ) : Ξ]]$.
\end{corollary}
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

\begin{lemma}[Algorithmic substitution equivalence becomes equality after normalization]
  \label{corollary:subst-subt-equiv-algorithmization}
  Suppose that $[[Θ ⊢ uσ1 : Ξ']]$ and $[[Θ ⊢ uσ2 : Ξ']]$ are algorithmic substitutions
  and $[[Ξ ⊆ Ξ']]$. 
  Then $[[Θ ⊢ uσ1 ≈ uσ2 : Ξ]] \iff [[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$.
\end{lemma}
\begin{proof}
  Follows immediately from \cref{lemma:subt-equiv-algorithmization}:
  \begin{itemize}
    \item [$\Rightarrow$]
      If $[[α̂± ∉ Ξ]]$, then $[[ [nf(uσ1)|Ξ]α̂± = [nf(uσ2)|Ξ]α̂± ]] = [[α̂±]]$ by 
      definition. 
      For any $[[α̂± ∊ Ξ]]$, 
      $[[ [nf(uσ1)|Ξ]α̂± = nf([uσ1]α̂±)]]$ and 
      $[[ [nf(uσ2)|Ξ]α̂± = nf([uσ2]α̂±) ]]$; 
      $[[Θ(α̂±) ⊢ [uσ1]α̂± ≈ [uσ2]α̂±]]$ implies
      $[[ nf([uσ1]α̂±) = nf([uσ2]α̂±) ]]$ by \cref{lemma:decl-equiv-algorithmization}.
    \item [$\Leftarrow$]
      If $[[α̂± ∊ Ξ]]$, then 
      $[[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$ implies
      $[[ nf([uσ1]α̂±) = nf([uσ2]α̂±) ]]$ by definition 
      of substitution restriction and normalization.
      In turn, $[[ nf([uσ1]α̂±) = nf([uσ2]α̂±) ]]$ means
       $[[Θ(α̂±) ⊢ [uσ1]α̂± ≈ [uσ2]α̂±]]$ by \cref{lemma:decl-equiv-algorithmization}.
  \end{itemize}
\end{proof}