### 2023-09-05 proofs/tex/wf-algo-lemmas.tex

\begin{lemma}[Soundness of algorithmic type well-formedness]
    \label{lemma:wf-algo-soundness}
    \hfill
  \begin{itemize}
    \item[$+$] if $[[Γ ; Ξ ⊢ uP]]$ then $[[fv(uP) ⊆ Γ]]$ and $[[uv(uP) ⊆ Ξ]]$;
    \item[$-$] if $[[Γ ; Ξ ⊢ uN]]$ then $[[fv(uN) ⊆ Γ]]$ and $[[uv(uN) ⊆ Ξ]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof is analogous to \cref{lemma:wf-soundness}.
  The additional base case is when $[[Γ ; Ξ ⊢ uP]]$ is derived by \ruleref{\ottdruleWFATPUVarLabel},
  and the symmetric negative case.
  In this case, $[[uP]] = [[α̂⁺]]$, and $[[uv(uP)]] = [[{α̂⁺} ⊆ Ξ]]$ by inversion; $[[fv(uP)]] = [[∅ ⊆ Γ]]$ vacuously.
\end{proof}

\begin{lemma}[Completeness of algorithmic type well-formedness]
  \label{lemma:wf-algo-ctxt-equiv}
  In the well-formedness judgment, only used variables matter:
  \begin{itemize}
  \item[$+$] if $[[Γ1 ∩ fv iP]] = [[Γ2 ∩ fv iP]]$
    and $[[Ξ1 ∩ uv uP]] = [[Ξ2 ∩ uv uP]]$ then
    $[[Γ1 ; Ξ1 ⊢ uP]] \iff [[Γ2 ; Ξ2 ⊢ uP]]$, and
  \item[$-$] if $[[Γ1 ∩ fv iN]] = [[Γ2 ∩ fv iN]]$
    and $[[Ξ1 ∩ uv uN]] = [[Ξ2 ∩ uv uN]]$ then
    $[[Γ1 ; Ξ1 ⊢ uN]] \iff [[Γ2 ; Ξ2 ⊢ uN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  By mutual structural induction on $[[uP]]$ and $[[uN]]$.
\end{proof}

\begin{lemma}[Variable algorithmization agrees with well-formedness]
  \label{lemma:var-algo-wf}
  \hfill
  \begin{itemize}
    \item[$+$]  $[[Γ, nas ⊢ iP]]$ implies $[[Γ; {nuas} ⊢ [nuas/nas]iP]]$;
    \item[$-$]  $[[Γ, nas ⊢ iN]]$ implies $[[Γ; {nuas} ⊢ [nuas/nas]iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof is a structural induction on $[[Γ, nas ⊢ iP]]$ and mutually, on $[[Γ, nas ⊢ iN]]$.
  Notice that the substitutions commute with all the constructors,
  providing the step of the induction. 
\end{proof}

\begin{lemma}[Variable dealgorithmization agrees with well-formedness]
  \label{lemma:var-dealgo-wf}
  \hfill
  \begin{itemize}
    \item[$+$]  $[[Γ; {nuas} ⊢ uP]]$ implies $[[Γ, nas ⊢ [nas/nuas]iP]]$;
    \item[$-$]  $[[Γ; {nuas} ⊢ uN]]$ implies $[[Γ, nas ⊢ [nas/nuas]iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  As for \cref{lemma:var-algo-wf}, the proof is a structural induction on 
  $[[Γ; {nuas} ⊢ uP]]$ and mutually, on $[[Γ; {nuas} ⊢ uN]]$.
\end{proof}

\begin{corollary}[Well-formedness Algorithmic Context Weakening]
  \label{lemma:wf-weakening-algo}
  Suppose that $[[Γ1 ⊆ Γ2]]$,
  and $[[Ξ1 ⊆ Ξ2]]$. Then
  \begin{itemize}
    \item[$+$] if $[[Γ1 ; Ξ1 ⊢ uP]]$ implies $[[Γ2 ; Ξ2 ⊢ uP]]$,
    \item [$-$] if $[[Γ1 ; Ξ1 ⊢ uN]]$ implies $[[Γ2 ; Ξ2 ⊢ uN]]$.
  \end{itemize}
\end{corollary}
\begin{proof}
  By \cref{lemma:wf-algo-soundness},
  $[[Γ1; Ξ1 ⊢ uP]]$ implies $[[fv(uP) ⊆ Γ1 ⊆ Γ2]]$
  and $[[uv(uP) ⊆ Ξ1 ⊆ Ξ2]]$,
  and thus, $[[fv(uP)]] = [[fv(uP) ∩ Γ1]] = [[fv(uP) ∩ Γ2]]$, 
  and $[[uv(uP)]] = [[uv(uP) ∩ Ξ1]] = [[uv(uP) ∩ Ξ2]]$.
  Then by \cref{lemma:wf-algo-ctxt-equiv}, $[[Γ2; Ξ2 ⊢ uP]]$. 
  The negative case is symmetric.
\end{proof}