### 2023-10-12 oopsla24/tex/03-declarative-system.tex
wip

The declarative system serves as a specification of the
type inference algorithm. It consists of two main parts:
the subtyping and the type inference.

\subsection{Subtyping}
It is represented by a set of inference rules shown in
\cref{fig:declarative-subtyping}.

\begin{figure}[h]
  \begin{multicols}{2}
    \ottdefnDOneNsub{}

    \ottdefnDOnePsup{}
  \end{multicols}
  \hfill\\
  \begin{multicols}{2}
    \ottdefnDOneNeq{}

    \ottdefnDOnePeq{}
  \end{multicols}
  \caption{Declarative Subtyping}
  \label{fig:declarative-subtyping}
\end{figure}

\paragraph{Quantifiers}  
Symmetric rules \ruleref{\ottdruleDOneForallLabel} and 
\ruleref{\ottdruleDOneExistsLabel} specify 
the subtyping between top-level quantified types.
Usually, the polymorphic subtyping is represented by two rules
introducing quantifiers to the left and to the right-hand side of the subtyping.
For conciseness of representation, we compose these rules into one.
First, our rule extends context $[[Γ]]$ with the quantified variables 
from the right-hand side ($[[pbs]]$ or $[[nbs]]$), 
as these variables must remain abstract.
Second, it verifies that the left-hand side quantifiers
($[[pas]]$ or $[[nas]]$) can be instantiated to continue subtyping recursively. 
The instantiation is modeled by substitution $[[σ]]$.
The notation $[[Γ2 ⊢ σ : Γ1]]$ specifies its domain and range.
For instance, $[[Γ, pbs ⊢ σ : {pas}]]$ means that 
$[[σ]]$ maps the variables from $[[pas]]$ to (positive) types
well-formed in $[[Γ, pbs]]$.
This way, application $[[ [σ]iN ]]$ instantiates (replaces) every
$[[αi⁻]]$ in $[[iN]]$ with $[[σ]]([[αi⁻]])$.

\paragraph{Invariant Shifts}
An important restriction that we put on the subtyping system is
that the subtyping on shifted types requires their equivalence,
as shown in \ruleref{\ottdruleDOneShiftDLabel} and
\ruleref{\ottdruleDOneShiftULabel}. Relaxing both of these
invariants make the system equivalent to \systemf, 
and thus, undecidable. 
However, after certain changes 
\ruleref{\ottdruleDOneShiftDLabel} can be relaxed to the covariant form,
 as we will discuss in \cref{todo}.

\paragraph{Functions}
Standardly, the subtyping of function types is covariant in the return type
and contravariant in the argument type.

\paragraph{Variables}
The subtyping of variables is defined reflexively,
which is enough to ensure the reflexivity of subtyping in general.
The algorithm will use the fact that the subtypes of a variable 
coincide with its supertypes, which however is not true for an
arbitrary type.

A property that we extensively use is that the subtyping is reflexive and transitive. 
Moreover, any two \emph{positive} types have the least upper bound, which makes the positive subtyping a semilattice.

\begin{property}[Subtyping is well-defined]
  The subtyping forms a preorder and is preserved under substitutions. 
\end{property}

\subsection{Equivalence and Normalization}

The subtyping relation forms a preorder on types,
and thus, it induces an equivalence relation \aka \emph{bicoercibility}.
The declarative specification of subtyping must be defined up to this equivalence.
Moreover, the algorithms we use must withstand changes in input types within the equivalence class.
To deal with this, we use normalization---a function that uniformly selects a representative of the equivalence class.

Using normalization gives us two benefits:
\begin{enumerate*}
  \item [(i)] we do not need to modify significantly standard operations such as unification to withstand non-trivial equivalence, and
  \item [(ii)] once the subtyping (and thus, the equivalence) changes, we only need to modify the normalization function, 
    while the rest of the algorithm remains the same.
\end{enumerate*}

\begin{figure}[h]
  \label{alg:type-nf}
  \hfill

  \begin{multicols}{2}
  \ottdefnNrmNNorm{}
  \columnbreak\\
  \ottdefnNrmPNorm{}
  \end{multicols}
  \caption{Type Normalization}
  \label{fig:type-nf}
\end{figure}

\subsection{Type Inference}

The declarative type inference system is shown in \cref{fig:declarative-inference}.
The positive typing judgment $[[Γ ; Φ ⊢ v : iP]]$ is read as 
``under the type context $[[Γ]]$ and variable context $[[Φ]]$,
the term $[[v]]$ is allowed to have the type $[[iP]]$'',
where $[[Φ]]$---the variable context---is defined standardly as
a set of pairs of the form $[[x : iP]]$. 
The negative typing judgment is read similarly.
To infer the resulting type of an application, the
applicative let rules refer to the \emph{Application typing} judgment.
It has form of $[[Γ ; Φ ⊢ iN ● args ⇒> iM]]$, 
which is read as ``under the type context $[[Γ]]$ and variable context $[[Φ]]$,
the application of a function of type $[[iN]]$ to the list of arguments $[[args]]$
is allowed to have the type $[[iM]]$''.

\begin{figure}[h]
  \ottdefnDTNInf{}

  \hfill

  \begin{multicols}{2}
  \ottdefnDTPInf{}
  \\
  \ottdefnDTSpinInf{}
  \end{multicols}
  \hfill

  \caption{Declarative Subtyping}
  \label{fig:declarative-inference}
\end{figure}

\paragraph{Non-bidirectionality}
  Notice that all the rules in our system are inferring.

\paragraph{Typing up to equivalence}
  As discussed in \cref{todo}, the subtyping, as a preorder, 
  induces a non-trivial equivalence relation on types. 
  The system must not distinguish between equivalent types,
  and thus, type inference must be defined up to equivalence. 
  For this purpose, we use rules \ruleref{\ottdruleDTPEquivLabel}  
  and \ruleref{\ottdruleDTNEquivLabel}.
  The application typing is defined up to equivalence as well,
  although it has no special rules for that.

\paragraph{Unpack}
  Rule \ruleref{\ottdruleDTUnpackLabel} types elimination of $[[∃]]$.
  First, it infers the normalized type of the existential package.
  The normalization is required to fix the order of the quantifying variables
  to bind them. After the bind, the rule infers the type of the body 
  and checks that it does not use the bound variables so that they do not 
  escape the scope.

\paragraph{Applicative Let Binders}
  Rules \ruleref{\ottdruleDTAppLetLabel} and \ruleref{\ottdruleDTAppLetAnnLabel}
  infer the type of the applicative let binders.
  Both of them infer the type of the head $[[v]]$ 
  and invoke the application typing to infer the type of the application 
  before recursing on the body of the let binder.
  The difference is that the former rule is for the 
  \emph{unannotated} let binder, and thus
  it requires the resulting type of application to be unique
  (up to equivalence), so that the type of the bound variable $[[x]]$
  is known before it is put into the context.
  The latter rule is for the \emph{annotated} binder,
  and thus, the type of the bound $[[x]]$ is given, 
  however, the rule must check that this type is a
  a supertype of the inferred type of the application. 
  This check is done by invoking the subtyping judgment
  $[[Γ ⊢ ↑iQ ≤ ↑iP]]$.
  This judgment is more restrictive than checking bare 
  $[[Γ ⊢ iP ≥ iQ]]$, however, it is necessary
  to make the algorithm complete as it allows us to preserve
  certain invariants (see \cref{todo}).
  In \cref{todo} we discuss how this restriction can be relaxed
  together with invariant shift subtyping.

\paragraph{Application of a polymorphic type $[[∀]]$}
  The complexity of the system is hidden in the rules, 
  whose output type is not immediately defined by their input
  and the output of their premises 
  (\aka not mode-correct \cite{dunfield2020:bidirectional}).
  In our typing system, such rule is \ruleref{\ottdruleDTForallAppLabel}: 
  the instantiation of the quantifying variables is
  taken out of thin air. The algorithm we present in \cref{todo} 
  delays this instantiation until more information about it 
  (in particular, typing constraints) is collected.

\paragraph{Application of an Arrow Type}
  Another important application rule is \ruleref{\ottdruleDTArrowAppLabel}.
  This is where the subtyping is used to check that the type of the argument
  is convertible to (a subtype of) the type of the function parameter.
  In the algorithm (\cref{todo}), this subtyping check will provide the constraints
  we need to resolve the delayed instantiations of the quantifying variables.

\paragraph{Annotations}
  Subtyping is also used by the annotation rules \ruleref{\ottdruleDTNAnnotLabel}
  and \ruleref{\ottdruleDTPAnnotLabel}. The annotation is only valid if the
  inferred type is a subtype of the annotation type.

\subsection{Properties of the Declarative System}

Now we present selected properties of the declarative system,
which are important for the correctness of the algorithm.

As we discussed in \cref{todo}, the equivalence induced by the subtyping is 
not trivial. To deal with it without modifying standard operations such as unification, 
we use normalization. This way, it is important for the normalization to be 
sound and complete \wrt the equivalence:

\begin{property}[Correctness of normalization]
  For $[[iN]]$ and $[[iM]]$ well-formed in $[[Γ]]$,
  $[[Γ ⊢ iN ≈ iM]]$ is equivalent to $[[nf(iN) = nf(iM)]]$.
\end{property}

However, the greatest lower bound of two positive types is not always defined.
Let us consider a positive type $[[iP]]$,
and two different supertypes $[[iP1]]$ and $[[iP2]]$
(\ie  $[[· ⊢ iP1 ≥ iP]]$, $[[· ⊢ iP2 ≥ iP]]$, and $[[nf(iP1) ≠ nf(iP2)]]$.
Let us also consider an arbitrary negative type $[[iN]]$.