### 2023-10-17 oopsla24/tex/01-introduction.tex
some more intro text

% How we should frame this:

Over the last half-century, there has been considerable work on developing type inference algorithms for programming languages,  mostly aimed at solving the problem of \emph{type polymorphism}.

That is, in pure polymorphic lambda calculus (system F), the polymorphic type $\forall a.\,A$ has a big lambda $\Lambda a.\,e$ as an introduction form, and an explicit type application $e\,[A]$ as an elimination form. This is an extremely simple and compact type system, whose rules fit on a single page, but whose semantics are sophisticated enough to model things like parametricity and representation independence. However, System F by itself is unwieldy as a programming language. The fact that the universal type $\forall a.\,A$ has explicit eliminations means that programs written using polymorphic types will need to be stuffed to the gills with type annotations explaining how and when to instantiate the quantifiers.

Therefore, most work on type inference has been aimed at handling type instantiations implicitly -- we want to be able to use a polymorphic function like $\mathsf{len} : \forall a.\, \mathsf{List}\,a \to \mathsf{int}$ at any concrete list type without explicitly instantiating the quantifier in $\mathsf{len}$'s type. That is, we want to write $\mathsf{len}\,[1,2,3]$ instead of writing $\mathsf{len}\,[\mathsf{int}]\,[1,2,3]$.

The most famous of the algorithms for solving these constraints is the Damas-Hindley-Milner algorithm. The idea is that type instantiation induces a subtype ordering: the type $\forall a.\,A$ is a subtype of all of its instantiations. So we wish to be free to use the same function $\mathsf{len}$ at many different types such as
$\mathsf{List}\,\mathsf{int} \to \mathsf{int}$,
$\mathsf{List}\,\mathsf{bool} \to \mathsf{int}$,
$\mathsf{List}\,(\mathsf{int \times bool}) \to \mathsf{int}$, and so on. However, the subtype relation is nondeterministic: it tells us that whenever we see a polymorphic type $\forall a.\, A$, we know it is a subtype of \emph{any} of its instantiations. To turn this into an algorithm, we have to actually make some choices, and DHM works by using \emph{unification}. Whenever we would have had to introduce a particular concrete type in the specification, the DHM algorithm introduces a unification variable, and incrementally instantiates this variable as more and more type constraints are observed.

However, the universal quantifier is not the only quantifier! Dual to the universal quantifer $\forall$ is the existential quantifier $\exists$. Even though existential types have an equal logical status to universal types, they have been studied much less frequently in the literature. The most widely-used algorithm for handling existentials is the algorithm of Odersky and Laufer. This algorithm treats existentials in a second-class fashion: they are not explicit connectives of the logic, but rather are tied to datatype declarations. As a result, packing and unpacking existential types is tied to constructor introductions and constructor eliminations in pattern matches.

\begin{itemize}
\item There has been very little work on type inference for existentials. 
\item The state of the art in languages like Ocaml and Haskell is the
  Odersky-L\"{a}ufer algorithm, which makes existentials a
  second-class construct tied to datatypes.
\item More recently, there's the existential crisis paper, which
  showed how to extend ML-style inference with some support for
  existentials. However, it is (a) predicative-only, (b) restricted
  for $\forall\exists$ types, and (c) does not have a declarative
  specification. (In fact, the authors comment on the need for such.)
\item In this paper, we lift all of these restrictions. We give a
  local type inference algorithm which supports first-class
  existentials, for impredicative types, with no restrictions on
  nested quantifers, and we have declarative specification we prove
  our algorithm sound and complete with respect to.
\item Developing the algorithm required us to break some of the
  fundamental invariants of HM-style typin, and as a result, our
  algorithm needs to mix unification and anti-unification, and works
  over a call-by-push-value metalanguage to control how quantifiers
  are instantiated and things like function arities.
\item Unfortunately, local type inference does not infer types anywhere
  nearly as comprehensively as full Damas-Milner type inference. We
  do guarantee that all type applications (for $\forall$-elimination) 
  and all packs (for $\exists$-introdution) are inferred, but functions
  and some let-bindings still need annotations on their arguments.
\item The original local type inference paper combined local type inference
  with bidirectional typechecking to minimize the number of needed
  annotations, but we show how existential types complicate the integration
  of bidirectionality with local type inference, and we explore the design space
  to show how the same scheme could be applied to work with different
  type systems. 
\end{itemize}

\nk{We need to explain what local type inference is, and how it is and isn't
  related to bidirectional typechecking.}