### 2023-11-13 proofs/tex/alg-neg-subtyping-statements.tex

\begin{restatable}[Negative Algorithmic Subtyping is Deterministic]{observation}{obsNegSubtDeterministic}
    \label{obs:neg-subt-deterministic}
    For fixed $[[Γ]]$, $[[Θ]]$, $[[iM]]$, and $[[uN]]$, 
    if $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$ 
    and $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC']]$
    then $[[SC]]$ = $[[SC']]$.
\end{restatable}

\begin{restatable}[Soundness of Negative Subtyping]{lemma}{lemNegSubtSoundness}
    \label{lemma:neg-subt-soundness}
    If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iM]]$, $[[Γ ; dom(Θ) ⊢ uN]]$ and 
    $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$, then 
    $[[Θ ⊢ SC : uv(uN)]]$ and 
    for any normalized $[[uσ]]$ such that $[[ Θ  ⊢ uσ : SC ]]$,
    $[[ Γ ⊢ [uσ]uN ≤ iM ]]$.
\end{restatable}

\begin{restatable}[Completeness of the Negative Subtyping]{lemma}{lemNegSubtCompleteness}
    \label{lemma:neg-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iM]]$, $[[Γ ; dom(Θ) ⊢ uN]]$,
    and $[[uN]]$ does not contain negative unification variables ($[[α̂⁻]] \notin [[uv uN]]$).
    Then for any $[[Θ ⊢ uσ : uv(uN)]]$ such that $[[Γ ⊢ [uσ]uN ≤ iM]]$,
    there exists $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$ and moreover, $[[ Θ ⊢ uσ : SC ]]$.
\end{restatable}