### 2023-11-13 proofs/tex/constraint-merge-statements.tex

\begin{restatable}[Positive Subtyping is Deterministic]{observation}{obsPosSubtDeterministic}
    \label{obs:pos-subt-deterministic}
    For fixed $[[Γ]]$, $[[Θ]]$, $[[uP]]$, and $[[iQ]]$,
    if $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ and $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC']]$
    then $[[SC]] = [[SC']]$.
\end{restatable}

\begin{restatable}[Constraint Entry Merge is Deterministic]{observation}{obsEntryMergeDeterministic}
    \label{obs:entry-merge-deterministic}
    For fixed $[[Γ]]$, $[[scE1]]$, $[[scE2]]$,
    if $[[Γ ⊢ scE1 & scE2 = scE]]$ and $[[Γ ⊢ scE1 & scE2 = scE']]$ 
    then $[[scE]] = [[scE']]$.
\end{restatable}

\begin{restatable}[Subtyping Constraint Merge is Deterministic]{observation}{obsSubtMergeDeterministic}
    \label{obs:subt-merge-deterministic}
    Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$
    If $[[Θ ⊢ SC1 & SC2 = SC]]$ and $[[Θ ⊢ SC1 & SC2 = SC']]$ are defined then 
    $[[SC]] = [[SC']]$.
\end{restatable}

\begin{restatable}[Soundness of Constraint Entry Merge]{lemma}{lemEntryMergeSoundness}
    \label{lemma:entry-merge-soundness}
    For a fixed context $[[Γ]]$,
    suppose that  $[[Γ ⊢ scE1]]$ and $[[Γ ⊢ scE2]]$. 
    If $[[Γ ⊢ scE1 & scE2 = scE]]$ is defined then
    \begin{enumerate}
        \item $[[Γ ⊢ scE]]$
        \item For any $[[Γ ⊢ iP]]$, $[[Γ ⊢ iP : scE]]$ implies $[[Γ ⊢ iP : scE1]]$ and $[[Γ ⊢ iP : scE2]]$ 
    \end{enumerate}
\end{restatable}

\begin{restatable}[Soundness of Constraint Merge]{lemma}{lemMergeSoundness} 
    \label{lemma:merge-soundness}
    Suppose that $[[Θ ⊢ SC1 : Ξ1]]$ and $[[Θ ⊢ SC2 : Ξ2]]$ 
    and $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
    Then 
    \begin{enumerate}
        \item $[[Θ ⊢ SC : Ξ1 ∪ Ξ2]]$,
        \item for any substitution $[[Θ ⊢ uσ : Ξ1 ∪ Ξ2]]$, 
            $[[ Θ ⊢ uσ : SC ]]$
            implies $[[ Θ ⊢ uσ : SC1 ]]$ and $[[ Θ ⊢ uσ : SC2 ]]$.
    \end{enumerate}
\end{restatable}

\begin{restatable}[Completeness of Constraint Entry Merge]{lemma}{lemEntryMergeCompleteness}
    \label{lemma:entry-merge-completeness}
    For a fixed context $[[Γ]]$,
    suppose that $[[Γ ⊢ scE1]]$ and $[[Γ ⊢ scE2]]$ are matching constraint entries.
    \begin{itemize}
        \item for a type $[[iP]]$ such that $[[Γ ⊢ iP : scE1]]$ and $[[Γ ⊢ iP : scE2]]$,
        $[[Γ ⊢ scE1 & scE2 = scE]]$ is defined and $[[Γ ⊢ iP : scE]]$.
        \item for a type $[[iN]]$ such that $[[Γ ⊢ iN : scE1]]$ and $[[Γ ⊢ iN : scE2]]$,
        $[[Γ ⊢ scE1 & scE2 = scE]]$ is defined and $[[Γ ⊢ iN : scE]]$.
    \end{itemize}
\end{restatable}

\begin{restatable}[Completeness of Constraint Merge]{lemma}{lemMergeCompleteness} 
    \label{lemma:merge-completeness}
    Suppose that $[[Θ ⊢ SC1 : Ξ1]]$ and $[[Θ ⊢ SC2 : Ξ2]]$.
    If there exists a substitution $[[Θ ⊢ uσ : Ξ1 ∪ Ξ2]]$ such that 
    $[[ Θ ⊢ uσ : SC1 ]]$ and $[[ Θ ⊢ uσ : SC2 ]]$
    then $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
\end{restatable}