### 2023-11-13 proofs/tex/constraint-sat-statements.tex

\begin{restatable}[Any constraint is satisfiable]{lemma}{lemmaConstraintSat}
    \label{lemma:constraint-sat}
    Suppose that $[[Θ ⊢ SC]]$
    and $[[Ξ]]$ is a set such that
    $[[dom(SC)]] \subseteq [[Ξ]] \subseteq [[dom(Θ)]]$.
    Then there exists 
    $[[uσ]]$ such that 
    $[[ Θ  ⊢ uσ : Ξ ]]$ and
    $[[ Θ  ⊢ uσ : SC ]]$.
\end{restatable}

\begin{restatable}[Constraint Entry Satisfaction is Stable under Equivalence]{lemma}{lemmaEntrySatEquiv}
    \label{lemma:entry-sat-equiv}
    \begin{itemize}
        \item [$-$] If $[[Γ ⊢ iN1 : scE]]$ and $[[Γ ⊢ iN1 ≈ iN2]]$ then $[[Γ ⊢ iN2 : scE]]$.
        \item [$+$] If $[[Γ ⊢ iP1 : scE]]$ and $[[Γ ⊢ iP1 ≈ iP2]]$ then $[[Γ ⊢ iP2 : scE]]$.
    \end{itemize}
\end{restatable}

\begin{restatable}[Constraint Satisfaction is stable under Equivalence]{corollary}{corollarySatEquiv}
    \label{corollary:sat-equiv}
    \hfill\\
    If $[[ Θ ⊢ uσ1 : SC ]]$ and $[[Θ ⊢ uσ1 ≈ uσ2 : dom(SC) ]]$ then $[[ Θ ⊢ uσ2 : SC ]]$;\\
    if $[[ Θ ⊢ uσ1 : UC ]]$ and $[[Θ ⊢ uσ1 ≈ uσ2 : dom(SC) ]]$ then $[[ Θ ⊢ uσ2 : UC ]]$.
\end{restatable}

\begin{restatable}[Normalization preserves Constraint Satisfaction]{corollary}{corollaryNfSat}
    \label{corollary:nf-sat}
    \hfill\\
    If $[[Θ ⊢ uσ : SC]]$ then $[[Θ  ⊢ nf(uσ) : SC]]$;\\
    if $[[Θ ⊢ uσ : UC]]$ then $[[Θ  ⊢ nf(uσ) : UC]]$.
\end{restatable}