### 2023-11-13 proofs/tex/equiv-alg-lemmas.tex

\lemmaWfEquivAlgo*
\begin{proof}
  The proof coincides with the proof of \cref{lemma:wf-equiv},
  and adds two cases for equating two positive or two negative algorithmic variables,
  which must be equal by inversion, and thus, 
  $[[Γ; Ξ ⊢ α̂±]] \iff [[Γ; Ξ  ⊢ α̂±]]$ holds trivially.
\end{proof}

\corollaryWfNfAlgo*
\begin{proof}
  Immediately from \cref{lemma:wf-equiv-algo,lemma:normalization-soundness-alg}.
\end{proof}

\corollaryNormSubstSigAlgo*
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

\corollarySubstSubtEquivAlg*
\begin{proof}
  Follows immediately from \cref{lemma:subt-equiv-algorithmization}:
  \begin{itemize}
    \item [$\Rightarrow$]
      If $[[α̂± ∉ Ξ]]$, then $[[ [nf(uσ1)|Ξ]α̂± = [nf(uσ2)|Ξ]α̂± ]] = [[α̂±]]$ by 
      definition. 
      For any $[[α̂± ∊ Ξ]]$, 
      $[[ [nf(uσ1)|Ξ]α̂± = nf([uσ1]α̂±)]]$ and 
      $[[ [nf(uσ2)|Ξ]α̂± = nf([uσ2]α̂±) ]]$; 
      $[[Θ(α̂±) ⊢ [uσ1]α̂± ≈ [uσ2]α̂±]]$ implies
      $[[ nf([uσ1]α̂±) = nf([uσ2]α̂±) ]]$ by \cref{lemma:decl-equiv-algorithmization}.
    \item [$\Leftarrow$]
      If $[[α̂± ∊ Ξ]]$, then 
      $[[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$ implies
      $[[ nf([uσ1]α̂±) = nf([uσ2]α̂±) ]]$ by definition 
      of substitution restriction and normalization.
      In turn, $[[ nf([uσ1]α̂±) = nf([uσ2]α̂±) ]]$ means
       $[[Θ(α̂±) ⊢ [uσ1]α̂± ≈ [uσ2]α̂±]]$ by \cref{lemma:decl-equiv-algorithmization}.
  \end{itemize}
\end{proof}