### 2023-11-13 proofs/tex/equiv-alg-statements.tex

\begin{restatable}[Algorithmic type well-formedness is invariant under equivalence]{lemma}{lemmaWfEquivAlgo}
    \label{lemma:wf-equiv-algo}
    Mutual subtyping implies declarative equivalence.
    \begin{itemize}
    \item[$+$] if $[[uP ≈ uQ]]$ then $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ uQ]]$,
    \item[$-$] if $[[uN ≈ uM]]$ then $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ uM]]$
    \end{itemize}
\end{restatable}

\begin{restatable}[Normalization preserves well-formedness of algorithmic types]{corollary}{corollaryWfNfAlgo}
    \label{corollary:wf-nf-algo}
    \hfill\\
    \begin{itemize}
        \item[$+$] $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ nf(uP)]]$,
        \item[$-$] $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ nf(uN)]]$
    \end{itemize}
\end{restatable}

\begin{restatable}[Normalization preserves the signature of the algorithmic substitution]{corollary}{corollaryNormSubstSigAlgo}
    \label{corollary:norm-subst-sig-algo}
    \hfill\\
    $[[Θ ⊢ uσ : Ξ]] \iff [[Θ ⊢ nf(uσ) : Ξ]]$,
    $[[Γ ⊢ uσ : Ξ]] \iff [[Γ ⊢ nf(uσ) : Ξ]]$.
\end{restatable}

\begin{restatable}[Algorithmic substitution equivalence becomes equality after normalization]{corollary}{corollarySubstSubtEquivAlg}
    \label{corollary:subst-subt-equiv-algorithmization}
    Suppose that $[[Θ ⊢ uσ1 : Ξ']]$ and $[[Θ ⊢ uσ2 : Ξ']]$ are algorithmic substitutions
    and $[[Ξ ⊆ Ξ']]$. 
    Then $[[Θ ⊢ uσ1 ≈ uσ2 : Ξ]] \iff [[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$.
\end{restatable}