### 2023-11-13 proofs/tex/norm-alg-lemmas.tex

\lemmaUVNF*
\begin{proof}
  By straightforward induction on $[[uN]]$ and mutually on $[[uP]]$, 
  similar to the proof of \cref{lemma:fv-nf}.
\end{proof}

\lemmaNormalizationSoundnessAlg*
\begin{proof}
  The proof coincides with the proof of \cref{lemma:normalization-soundness}.
\end{proof}