### 2023-11-13 proofs/tex/ord-statements.tex

\begin{restatable}[Ordering is deterministic]{observation}{obsOrdDeterministic}
    \label{obs:ord-deterministic}
    If $[[ord varset in iN = ordVars1]]$ and $[[ord varset in iN = ordVars2]]$ then $[[ordVars1 = ordVars2]]$.
    If $[[ord varset in iP = ordVars1]]$ and $[[ord varset in iP = ordVars2]]$ then $[[ordVars1 = ordVars2]]$.
    This way, we can use $[[ord varset in iN]]$ and as a function on $[[iN]]$,
    and $[[ord varset in iP]]$ as a function on $[[iP]]$.
\end{restatable}

\begin{restatable}[Soundness of variable ordering]{lemma}{lemOrdSoundness}
  \label{lemma:ord-soundness}
  Variable ordering extracts used free variables.
  \begin{itemize}
    \item[$-$] $[[ {ord varset in iN} ]] = [[varset ∩ fv iN]]$ (as sets)
    \item[$+$] $[[ {ord varset in iP} ]] = [[varset ∩ fv iP]]$ (as sets)
  \end{itemize}
\end{restatable}

\begin{restatable}[Additivity of ordering]{corollary}{corOrdAdditivity}
  \label{corollary:ord-additivity}
  Variable ordering is additive (in terms of set union) with respect to its first argument.
  \begin{itemize}
    \item[$-$] $[[ {ord (varset1 ∪ varset2) in iN} 
                = 
                {ord varset1 in iN} ∪ {ord varset2 in iN}]]$ (as sets)
    \item[$+$] $[[{ord (varset1 ∪ varset2) in iP}
                =
                {ord varset1 in iP} ∪ {ord varset2 in iP}]]$ (as sets)

  \end{itemize}
\end{restatable}

\begin{restatable}[Weakening of ordering]{lemma}{lemOrdWeakening}
  \label{corollary:ord-weakening}
  Only used variables matter in the first argument of the ordering,
  \begin{itemize}
    \item[$-$] $[[ ord (varset ∩ fv iN) in iN ]] = [[ ord varset in iN ]]$
    \item[$+$] $[[ ord (varset ∩ fv iP) in iP ]] = [[ ord varset in iP ]]$
  \end{itemize}
\end{restatable}

\begin{restatable}[Idempotency of ordering]{corollary}{corOrdIdemp}
  \label{corollary:ord-idemp}
  \hfill
  \begin{itemize}
    \item[$-$] If $[[ ord varset in iN = ordVars ]]$ then 
      $[[ ord {ordVars} in iN = ordVars ]]$,
    \item[$+$] If $[[ ord varset in iP = ordVars ]]$ then 
      $[[ ord {ordVars} in iP = ordVars ]]$;
  \end{itemize}
\end{restatable}

\begin{restatable}[Distributivity of renaming over variable ordering]{lemma}{lemDistrMuOrd}
  \label{lemma:distr-mu-ord}
  Suppose that $\mu$ is a bijection between two sets of variables
  $\mu : A \leftrightarrow B$.

  \begin{itemize}
  \item[$-$]
    If $[[mu]]$ is collision-free on $[[varset]]$ and $[[fv iN]]$ then
    $[[ [mu] (ord varset in iN) ]] = [[ord ([mu] varset) in [mu] iN ]]$
  \item[$+$]
    If $[[mu]]$ is collision-free on $[[varset]]$ and $[[fv iP]]$ then
    $[[ [mu] (ord varset in iP) ]] = [[ord ([mu] varset) in [mu] iP ]]$
  \end{itemize}
\end{restatable}

\begin{restatable}[Ordering is not affected by independent substitutions]{lemma}{lemOrdSigma}
    \label{lemma:ord-sigma}
    Suppose that $[[Γ2 ⊢ σ : Γ1]]$, i.e. $[[σ]]$ maps variables from $[[Γ1]]$ into types
    taking free variables from $[[Γ2]]$, and $[[varset]]$ is a set of variables
    disjoint with both $[[Γ1]]$ and $[[Γ2]]$, 
    $[[iN]]$ and $[[iP]]$ are types. Then
        \begin{itemize}
    \item[$-$] $[[ ord varset in [σ]iN ]] = [[ord varset in iN ]]$
    \item[$+$] $[[ ord varset in [σ]iP ]] = [[ord varset in iP ]]$
    \end{itemize}
\end{restatable}

\begin{restatable}[Completeness of variable ordering]{lemma}{lemOrdCompleteness}
    \label{lemma:ord-completeness}
    Variable ordering is invariant under equivalence. For arbitrary $[[varset]]$,
     \begin{itemize}
    \item[$-$] If $[[iN ≈ iM]]$ then $[[ord varset in iN]] = [[ord varset in iM]]$ (as lists)
    \item[$+$] If $[[iP ≈ iQ]]$ then $[[ord varset in iP]] = [[ord varset in iQ]]$ (as lists)
    \end{itemize}
\end{restatable}