### 2023-11-13 proofs/tex/subst-statements.tex

\begin{restatable}[Substitution strengthening]{lemma}{lemmaSubstRestrFv}
    \label{lemma:subst-restr-fv}
    Restricting the substitution to the free variables of the
    substitution subject does not affect the result.
    Suppose that $[[σ]]$ is a substitution,
    $[[iP]]$ and $[[iN]]$ are types. Then
    \begin{itemize}
        \item[$+$] $[[ [σ]iP ]] = [[ [σ|fv iP]iP ]]$,
        \item[$-$] $[[ [σ]iN ]] = [[ [σ|fv iN]iN ]]$
    \end{itemize}
\end{restatable}

\begin{restatable}[Signature of a restricted substitution]{lemma}{lemmaSubstRestrSig}
    \label{lemma:subst-restr-sig}
    If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2 ⊢ σ|varset :  Γ1 ∩ varset]]$.
\end{restatable}

\begin{restatable}{lemma}{lemmaSubstDisjoint}
    \label{lemma:subst-disjoint}
    Suppose that $[[σ]]$ is a substitution with signature $[[Γ2 ⊢ σ : Γ1]]$. 
    Then if $[[varset]]$ is disjoint from $[[Γ1]]$, 
    then $[[ σ|varset]] = [[ id ]]$. 
\end{restatable}

\begin{restatable}[Application of a disjoint substitution]{corollary}{corollarySubstDisj}
    \label{corollary:subst-disj}
    Suppose that $[[σ]]$ is a substitution with signature $[[Γ2 ⊢ σ : Γ1]]$. Then
    \begin{itemize}
        \item [$+$] if $[[Γ1 ∩ fv(iQ) = ∅]]$ then $[[ [σ] iQ = iQ]]$;
        \item [$-$] if $[[Γ1 ∩ fv(iN) = ∅]]$ then $[[ [σ] iN = iN]]$.
    \end{itemize}
\end{restatable}

\begin{restatable}[Substitution range weakening]{lemma}{lemmaSubstRangeWeak}
    \label{lemma:subst-range-weakening}
    Suppose that $[[Γ2 ⊆ Γ'2]]$ are contexts and 
    $[[σ]]$ is a substitution.
    Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ'2 ⊢ σ : Γ1]]$.
\end{restatable}

\begin{restatable}[Subsitutions Equivalent on Free Variables]{lemma}{lemmaSubstEquiv}
    \label{lemma:subst-equiv}
    Suppose that $[[Γ' ⊆ Γ]]$,
    $[[σ1]]$ and $[[σ2]]$ are substitutions of signature $[[Γ ⊢ σi : Γ']]$.
    Then 
    \begin{enumerate}
        \item [$+$] for a type $[[Γ ⊢ iP]]$, if $[[Γ ⊢ [σ1]iP ≈ [σ2]iP]]$ then 
        $[[Γ ⊢ σ1 ≈ σ2 : fv iP ∩ Γ']]$;
        \item [$-$] for a type $[[Γ ⊢ iN]]$, if $[[Γ ⊢ [σ1]iN ≈ [σ2]iN]]$ then
        $[[Γ ⊢ σ1 ≈ σ2 : fv iN ∩ Γ']]$.
    \end{enumerate}
\end{restatable}

\begin{restatable}[Substitution composition well-formedness]{lemma}{lemmaSubstCompWell}
    \label{lemma:subst-comp-well}
    If $[[Γ'1 ⊢ σ1 : Γ1]]$ and $[[Γ'2 ⊢ σ2 : Γ2]]$,
    then $[[Γ'1, Γ'2 ⊢ σ2 ○ σ1 : Γ1, Γ2]]$.
\end{restatable}

\begin{restatable}[Substitution monadic composition well-formedness]{lemma}{lemmaSubstMonadCompWell}
    \label{lemma:subst-monad-composition-wf}
    If $[[Γ'1 ⊢ σ1 : Γ1]]$ and $[[Γ'2 ⊢ σ2 : Γ2]]$,
    then $[[Γ'2 ⊢ σ2 <=< σ1 : Γ1]]$.
\end{restatable}

\begin{restatable}[Substitution composition]{lemma}{lemmaSubstComp}
    \label{lemma:subst-composition}
        If $[[Γ'1 ⊢ σ1 : Γ1]]$, $[[Γ'2 ⊢ σ2 : Γ2]]$, 
        $[[Γ1 ∩ Γ'2 = ∅ ]]$ and $[[ Γ1 ∩ Γ2 = ∅ ]]$ then 
        $[[ σ2 ○ σ1 ]] = [[ (σ2 <=< σ1) ○ σ2 ]]$.
\end{restatable}

\begin{restatable}[Substitution composition commutativity]{corollary}{corollarySubstCompCumm}
    \label{corollary:subst-composition-commutativity}
    If $[[Γ'1 ⊢ σ1 : Γ1]]$, $[[Γ'2 ⊢ σ2 : Γ2]]$, and
    $[[ Γ1 ∩ Γ2 = ∅ ]]$, 
    $[[ Γ1 ∩ Γ'2 = ∅ ]]$, and
    $[[ Γ'1 ∩ Γ2 = ∅ ]]$ then 
    $[[ σ2 ○ σ1 ]] = [[ σ1 ○ σ2 ]]$.
\end{restatable}

\begin{restatable}[Substitution domain weakening]{lemma}{lemmaSubstDomWeak}
    \label{lemma:subst-domain-weakening}
    If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2, Γ' ⊢ σ : Γ1, Γ']]$
\end{restatable}

\begin{restatable}[Free variables after substitution]{lemma}{lemmaSubstFv}
    \label{lemma:subst-fv}
    Suppose that $[[Γ2 ⊢ σ : Γ1]]$, then
    \begin{enumerate}
        \item [$+$] for a type $[[iP]]$,
            the free variables of $[[ [σ]iP ]]$ are bounded in the following way:
         $[[fv(iP) \ Γ1 ⊆ fv([σ]iP) ⊆ (fv(iP) \ Γ1) ∪ Γ2]]$;
        \item [$-$] for a type $[[iN]]$, 
            the free variables of $[[ [σ]iP ]]$ are bounded in the following way:
            $[[fv(iN) \ Γ1 ⊆ fv([σ]iN) ⊆ (fv(iN) \ Γ1) ∪ Γ2]]$.
    \end{enumerate}
\end{restatable}

\begin{restatable}[Free variables of a variable image]{lemma}{lemmaSubstFvImage}
    \label{lemma:subst-fv-image}
    Suppose that $[[σ]]$ is an arbitrary substitution, 
    Then 
    \begin{itemize}
        \item [$+$] if $[[α± ∊ fv(iP)]]$ then $[[ fv([σ]α±) ⊆ fv([σ]iP) ]]$,
        \item [$-$] if $[[α± ∊ fv(iN)]]$ then $[[ fv([σ]α±) ⊆ fv([σ]iN) ]]$.
    \end{itemize}
\end{restatable}