### 2023-11-13 proofs/tex/wf-algo-lemmas.tex

\lemmaWfAlgoSoundness*
\begin{proof}
  The proof is analogous to \cref{lemma:wf-soundness}.
  The additional base case is when $[[Γ ; Ξ ⊢ uP]]$ is derived by \ruleref{\ottdruleWFATPUVarLabel},
  and the symmetric negative case.
  In this case, $[[uP]] = [[α̂⁺]]$, and $[[uv(uP)]] = [[{α̂⁺} ⊆ Ξ]]$ by inversion; $[[fv(uP)]] = [[∅ ⊆ Γ]]$ vacuously.
\end{proof}

\lemmaWfAlgoCtxtEquiv*
\begin{proof}
  By mutual structural induction on $[[uP]]$ and $[[uN]]$.
\end{proof}

\lemmaVarAlgoWf*
\begin{proof}
  The proof is a structural induction on $[[Γ, nas ⊢ iP]]$ and mutually, on $[[Γ, nas ⊢ iN]]$.
  Notice that the substitutions commute with all the constructors,
  providing the step of the induction. 
\end{proof}

\lemmaVarDealgoWf*
\begin{proof}
  As for \cref{lemma:var-algo-wf}, the proof is a structural induction on 
  $[[Γ; {nuas} ⊢ uP]]$ and mutually, on $[[Γ; {nuas} ⊢ uN]]$.
\end{proof}

\corollaryWfWeakeningAlgo*
\begin{proof}
  By \cref{lemma:wf-algo-soundness},
  $[[Γ1; Ξ1 ⊢ uP]]$ implies $[[fv(uP) ⊆ Γ1 ⊆ Γ2]]$
  and $[[uv(uP) ⊆ Ξ1 ⊆ Ξ2]]$,
  and thus, $[[fv(uP)]] = [[fv(uP) ∩ Γ1]] = [[fv(uP) ∩ Γ2]]$, 
  and $[[uv(uP)]] = [[uv(uP) ∩ Ξ1]] = [[uv(uP) ∩ Ξ2]]$.
  Then by \cref{lemma:wf-algo-ctxt-equiv}, $[[Γ2; Ξ2 ⊢ uP]]$. 
  The negative case is symmetric.
\end{proof}