### 2023-11-29 oopsla24/tex/06-extensions.tex
wip: extensions

\label{sec:extensions}

\subsection{Explicit Type Application}

In our system, all type applications are inferred implicitly: the algorithm
automatically instantiates the variables abstracted by $\forall$. 
The implicit type application can be added to the declarative system by the following rule:
$$\ottdruleDTTypeApp{}$$

However, this rule alone would cause ambiguity.
The declarative system does not fix the order of the quantifiers, 
which means that $[[∀α⁺.∀β⁺.iN]]$ and $[[∀β⁺.∀α⁺.iN]]$ can be inferred 
as a type of $[[c]]$ interchangeably. But then explicit instantiation
$([[ c[iP] ]])$ would be ambiguous, as it is unclear whether $[[α⁺]]$ or
$[[β⁺]]$ should be instantiated with $[[iP]]$.

\paragraph{Solution 1: Declarative Normalization}
One way to resolve this ambiguity is to fix the order of quantifiers.  
The algorithm already performs the ordering of the quantifiers in the normalization procedure.
This way, we could require the inferred type to be normalized
to specify the order of the quantifiers:
$$\ottdruleDTTypeAppOrdered{}$$

The drawback of this approach is that it 
would cause the `leakage' of the internal algorithmic 
concept of type normalization into the `surface' declarative 
system.

\paragraph{Solution 2: Elementary Type Inference}
An alternative approach to provide explicit type application
 was proposed by \cite{zhao22:elementary}.
In this work, the subtyping relation is restricted 
in such a way that 
$[[∀α⁺.∀β⁺.iN]]$ and $[[∀β⁺.∀α⁺.iN]]$ are \emph{not} mutual subtypes 
(as long as $[[α⁺ ∊ fv(iN)]]$ and $[[β⁺ ∊ fv(iN)]]$).
It implies that the order of the quantifiers of the inferred type is unique,
and thus, the explicit type application is unambiguous.

These restrictions can be incorporated into our system by
replacing the polymorphic subtyping rules
\ruleref{\ottdruleDOneForallLabel} and \ruleref{\ottdruleDOneExistsLabel}
with the following stronger versions:
\vspace{-\baselineskip}
\begin{multicols}{2}
$$ \ottdruleDOneEForallR{} $$\\
$$ \ottdruleDOneEForallL{} $$
\end{multicols}

\begin{multicols}{2}
$$ \ottdruleDOneEExistsR{} $$\\
$$ \ottdruleDOneEExistsL{} $$
\end{multicols}
\vspace{\baselineskip}
From the perspective of mutual subtyping,
these changes fix the order of the quantifiers for an equivalence class. 
Moreover, the equivalence degenerates to equality (alpha-equivalence).

To accommodate these changes in the \emph{algorithm}, 
it suffices to 
    (i) replace the normalization procedure with identity: $[[nf(iN)]] \defeq [[iN]]$, $[[nf(iP)]] \defeq [[iP]]$,
    (ii) modify the least upper bound polymorphic rule \ruleref{\ottdruleLUBExistsLabel} in the following way:
        $$\ottdruleLUBEExists{}$$ 
    and 
    (iii) replace the subtyping polymorphic rule \ruleref{\ottdruleAForallLabel} 
    by the following pair of rules:
    \begin{multicols}{2}
            $$\ottdruleAForallR{}$$\\
            $$\ottdruleAForallL{}$$
    \end{multicols}

\subsection{Bounded Quantification}

It is possible to smoothly extend the type system with bounded quantifiers.
In particular, we can add lower bounds to polymorphic $[[∀]]$-quantifiers:
$[[∀(pas ≥ iPs). iN]]$ with the expected subtyping specification:

$$ ??? $$

%   Γ, pbs ⊢ σ : {pas}    Γ, pbs ⊢ [σ] iN ≤ iM
%   ---------------------------------------- :: Forall
%   Γ ⊢ ∀(pas ≥ iPs). iN ≤ ∀(pbs ≥ iQs).iM

\subsection{Bidirectionalization and Invariants Relaxation}

The algorithm we provide requires that all lambda functions are annotated. To
augment the system's expressiveness and lessen this requirement, we can employ
bidirectionalization.

However, this approach will disrupt the crucial invariants of the system and
necessitate a more sophisticated constraint solver. Specifically, the generated
constraints may contain algorithmic variables on both sides of the subtyping
relation. If the constraint set only imposes an equivalence restriction, this
task aligns with pattern-unification. Nevertheless, the subtyping restrictions
render the problem undecidable. For instance, ...