### 2024-01-23 oopsla24/tex/07-conclusion.tex

We have presented a type inference algorithm for an impredicative polymorphic
lambda calculus with existential types and subtyping.
The system is designed in call-by-value paradigm, that allowed us to 
restrict it to a decidable fragment of the language,
that we describe declaratively. We presented the inference algorithm
and proved its soundness and completeness with respect to the specification.
The algorithm combines unification---a standard type inference technique---with
anti-unification---which has not been used in type inference before.