### 2024-04-15 oopsla24/tex/01-introduction.tex
Edit Introduction

Over the last half-century, there has been considerable work on developing type
inference algorithms for programming languages,  mostly aimed at solving the
problem of \emph{type polymorphism}.

That is, in pure polymorphic lambda
calculus---\systemf\cite{girard-system-f,jcr-system-f}---the polymorphic type
$[[∀a.fA]]$ has a big lambda $[[Λa.fe]]$ as an introduction form, and an
explicit type application $[[ fe[fA] ]]$ as an elimination form. This is an
extremely simple and compact type system, whose rules fit on a single page, but
whose semantics are sophisticated enough to model things like parametricity and
representation independence. However, System F by itself is unwieldy as a
programming language. The fact that the universal type $[[∀a.fA]]$ has explicit
eliminations means that programs written using polymorphic types will need to be
stuffed to the gills with type annotations explaining how and when to
instantiate the quantifiers.

Therefore, most work on type inference has been aimed at handling type
instantiations implicitly---we want to be able to use a polymorphic function
like $[[len : ∀a.List a → Int]]$ at any concrete list type without explicitly
instantiating the quantifier in $[[len]]$'s type. That is, we want to write
$[[ len [one,two,three] ]]$ instead of writing
$[[ len [Int] [one,two,three] ]]$.

The most famous of the algorithms for solving these constraints is the
Damas-Hindley-Milner (HDM) algorithm
\cite{hindley69:principal,milner78:theory,damas82:principal}. The idea is that
type instantiation induces a subtype ordering: the type $[[∀a.fA]]$ is a subtype
of all of its instantiations. So we wish to be free to use the same function
$[[len]]$ at many different types such as 
$[[List Int → Int]]$, $[[List Bool → Int]]$,
$[[List (Int × Bool) → Int]]$, and so on.
However, the subtype relation is nondeterministic: it tells us that whenever we
see a polymorphic type $[[∀a.fA]]$, we know it is a subtype of \emph{any}
of its instantiations. To turn this into an algorithm, we have to actually make
some choices, and DHM works by using \emph{unification}. Whenever we would have
had to introduce a particular concrete type in the specification, the DHM
algorithm introduces a unification variable, and incrementally instantiates this
variable as more and more type constraints are observed.

However, the universal quantifier is not the only quantifier! Dual to the
universal quantifier $\forall$ is the existential quantifier $\exists$. Even
though existential types have an equal logical status to universal types, they
have been studied much less frequently in the literature. The most widely-used
algorithm for handling existentials is the algorithm of Odersky and Laufer
\cite{laufer94:polymorphic}. This algorithm treats existentials in a
second-class fashion: they are not explicit connectives of the logic but rather
are tied to datatype declarations. As a result, packing and unpacking
existential types is tied to constructor introductions and constructor
eliminations in pattern matches. This allows Damas-Milner inference to continue
to work almost unchanged, but it does come at the cost of losing first-class
existentials and also of restricting the witness types to monomorphic types.

There has been a limited amount of work on support for existential types as a
first-class concept. In an unpublished draft, \citet{leijen06:first-class}
studied an extension of Damas-Milner inference in which type schemes contain
alternating universal and existential quantifiers. Quantifiers still range over
monotypes, and higher-rank polymorphism is not permitted. More recently,
\citet{dunfield16:existential} studied type inference for existential types in
the context of GADT type inference, which, while still predicative, supported
higher-rank types (i.e., quantifiers can occur anywhere inside of a type
scheme). \citet{eisenberg21:existential} propose a system which only permits
types of the form forall-exists, but which permits projective elimination in the
style of ML modules.

All of these papers are restricted to \emph{predicative} quantification, where
quantifiers can only be instantiated with monotypes (i.e., types without any
occurences of quantifiers). However, existential types in full System F are
\emph{impredicative}---that is, quantifiers can be instantiated with arbitrary
types, specifically including types containing quantifiers.

Historically, inference for impredicative quantification has been neglected, due
to results by \citet{tiuryn-urzczyn-96} and \citet{chrzaszcz-98} which show that
not only is full type inference for \systemf undecidable, but even that the
subtyping relation induced by type instantiation is undecidable. However, in
recent years, interest in impredicative inference has revived (for example, the
work of \citet{serrano-2020}), with a focus on avoiding the undecidability
barriers by doing \emph{partial} type inference. That is, we no longer try to do
full type inference, but rather accept that the programmer will need to write
annotations in cases where inference would be too difficult. Unfortunately, it
is often difficult to give a specification for what the partial algorithm
does--for example, \citet{eisenberg21:existential} observe that their algorithm
lacks a declarative specification, and explain why existential types make this
particularly difficult to do.

One especially well-behaved form of partial type inference is \emph{local type
inference}, introduced by \citet{pierce2000:local}. In this approach,
quantifiers in a polymorphic function are instantiated using only the type
information available in the arguments at each call site. While this infers
fewer quantifiers than global algorithms such as Damas-Milner can, it has the
twin benefits that it is easy to give a mathematical specification to, and that
failures of type inference are easily explained in terms of local features of
the call site. In fact, many production programming languages such as C\# and
Java use a form of local type inference, due to the ease of extending this style
of inference to support OO features.

In this paper, we extend the local type inference algorithm to a language with
both universal and existential quantifiers, which can both be instantiated
impredicatively. This combination of features broke a number of the invariants
which traditional type inference algorithms depend on, and required us to invent
new algorithms which combine both unification and (surprisingly)
anti-unification.

To research the type inference in impredicative polymorphic systems
in the presence of existential types, we make the following \emph{contributions}.

\paragraph{Type Inference Algorithm} We give a local type inference algorithm
    which supports both first-class existential and universal quantifiers, both
    of which can be instantiated impredicatively. To evade the undecidability
    results surrounding type inference for \systemf, we work in a variant of
    \CBPV \cite{levy2006:cbpv}, which lets us formulate a subtyping relation
    which is still decidable.

\paragraph{Declarative Specification}
    We give a declarative type system (again, based on \CBPV) to
    serve as a specification of our algorithm, and we prove our algorithm is
    sound and complete with respect to the declarative type system. The
    specification makes it easy to see that all type applications (for
    $\forall$-elimination) and all packs (for $\exists$-introdution) are
    inferred.

\paragraph{Employment of the Anti-unification}
    The introduction of impredicative existentials breaks some of the
    fundamental invariants of HM-style type inference. As a result, it needs to
    mix unification with \emph{anti-unification}---a dual operation which has
    been used in the literature on term rewriting systems, but which has never
    appeared in the context of polymorphic type inference.

\paragraph{Extrapolation of the Inference Framework to Other Type Systems}
    We explore the design space further to show how the
    same type inference scheme could be applied to
    work with different type systems. In particular,
    \begin{enumerate*}
      \item[(i)] in combination with \emph{bidirectional} typechecking---
        the approach used in the original local type inference paper of \citet{pierce2000:local}
        that minimizes the number of needed annotations;
      \item[(ii)] in systems with different subtyping rules, such as  
        the one used in \citet{zhao22:elementary}---a system that
        permits explicit type applications;
      \item[(iii)] in systems with \emph{bounded} polymorphic quantifiers.
        That is, one can only instantiate a quantifier with a 
        type that is a subtype of a given bound.
    \end{enumerate*}