### 2024-04-30 oopsla24/tex/02-overview.tex

\label{sec:overview}

In the presence of impredicative polymorphism, the standard \systemf has
undecidable subtyping and type inference. It means that any inference algorithm
is incomplete: it only infers the types for a \emph{fragment} of \systemf.

To tackle this issue, we restrict \systemf by employing the ideas from the \CBPV
system \cite{levy2006:cbpv}. We introduce a new language, \fexists, which
stratifies the syntax of \systemf into positive and negative parts. It provides
the structure that allows us to restrict the typing specification to make it
decidable.

Every term and type fo \fexists has either \emph{positive} or \emph{negative}
polarity. The type variables are annotated with their polarities (e.g. $[[α⁺]]$
or $[[β⁻]]$), and the types can change polarity via the shift operators $[[↑]]$
(positive to negative) and $[[↓]]$ (negative to positive).

Positive expressions correspond to \emph{values} in the \CBPV system, and
negative expressions correspond to \emph{computations}. The difference between
the computations and values is intuitively described in the motto of \CBPV: ``a
value is, a computation does''. In particular, an argument of a function is
always a value (\ie positive), and the resulting type, as well as the function
itself, is a computation (\ie negative). Analogously, the polymorphic
$\forall$ quantifies a computation over positive types.

\subsection{Examples}

Generally, any term and type of \systemf can be embedded into \fexists in
multiple ways. This is because at any given point, one can opt for either the
positive or negative polarity of the subterm. We will delve into the consistent
term-level and type-level embeddings (\ie preserving typing) in more detail in
\cref{sec:rel-to-systemf}. \Cref{fig:polarization-examples} showcases the
polarized versions of the standard \systemf terms. Next, we will discuss several
examples (using the terms from \cref{fig:polarization-examples}) to illustrate
what can and cannot be inferred in \fexists.

\begin{figure}[h]
  \begin{align*}
    [[map]] &: [[↓(∀α⁺.∀β⁺.↓(α⁺ → ↑β⁺) → List α⁺ → ↑List β⁺)]] \\
    [[len]] &: [[↓(∀α⁺.List α⁺ → ↑Int)]] \\
    [[choose]] &: [[↓(∀α⁺.α⁺ → α⁺ → ↑α⁺)]] \\
    [[id]] &: [[↓(∀α⁺.α⁺ → ↑α⁺)]] \\
    [[auto]] &: [[↓(↓(∀α⁺.α⁺ → ↑α⁺) → (∀α⁺.α⁺ → ↑α⁺))]]
  \end{align*}
  \caption{Polarization of standard \systemf terms}
  \label{fig:polarization-examples}
\end{figure}

\paragraph{Application of a Polymorphic Function}
  A polymorphic function can be applied to a value of a concrete type. In this
  case, the quantified variables are instantiated to make the application
  type-check. For instance, the polymorphic $[[len]]$ can be applied to a list
  of any positive type, and the resulting type will be inferred by the
  algorithm:
  $$[[· ; Φ ⊢ let x = len([one,two,three]); return x : ↑Int]]$$

\paragraph{Polymorphic Self-application}
  Unlike predicative systems, \fexists permits self-application of polymorphic
  functions. For instance, $[[id]]$ can be applied to itself resulting in a
  suspended polymorphic computation of type $[[↓(∀α⁺.α⁺ → ↑α⁺)]]$, which, in
  turn, can be applied to itself:
  $$[[· ; Φ ⊢ let x = id(id); let y = x(x); return y : ↑↓(∀α⁺.α⁺ → ↑α⁺)]]$$

\paragraph{Non-example: Polymorphic Arguments}
The focus of our system is the \emph{local} type inference
\cite{pierce2000:local}, which has certain limitations. In particular, the
instantiation does not happen when the polymorphic term is in the argument
position. For example, one could expect the following to be inferrable:
\mbox{$[[· ; Φ ⊢ let x = map(id, [two, three, nine]); return x : ↑ List Int]]$}.
However, to infer this, $[[id]]$ must be instantiated to \mbox{$[[Int → ↑Int]]$}
which requires the information to be propagated from the neighboring branch of
the syntax tree (\mbox{$[[ [two, three, nine] ]]$}), \ie bypass the locality.

The annotated version of the same term infers the type successfully, 
since all the polymorphic variables of the arguments are instantiated:
$$[[· ; Φ ⊢ let x = map((id : ↓(Int → ↑Int)), [two, three, nine]); return x : ↑ List Int]]$$

\paragraph{Inferring the Common Supertype}
  The function $[[choose]]$ takes two arguments of the same
  polymorphic type. To apply it to two values of different (concrete) types, 
  the inference algorithm must find the minimal type they can be coerced 
  to---the least common supertype.

  Let us assume that in $[[Φ]]$, we have two identity-like functions 
  \mbox{$[[id_iM]] : [[↓(↓iM → iM)]]$} and \mbox{$[[id_iN]] : [[↓(↓iN → iN)]]$}.
  Then their least common supertype---the existential $[[∃γ⁻.↓(↓γ⁻ → γ⁻)]]$
  is inferrable:
  $$[[· ; Φ ⊢ let x = choose (id_iN, id_iM); return x : ↑ ∃γ⁻.↓(↓γ⁻ → γ⁻)]]$$

\paragraph{Inferring a general type}
  As mentioned before, the local inference does not instantiate the polymorphic
  types in the argument position.
  Because of that, the application 
  \mbox{$[[let x = choose (id, auto); return x ]]$} cannot infer 
  $[[↑↓(↓(∀α⁺.α⁺ → ↑α⁺) → (∀α⁺.α⁺ → ↑α⁺))]]$.
  However, the inference succeeds and infers a less informative type:
  $$[[· ; Φ ⊢ let x = choose (id, auto); return x : ↑ ∃γ⁻.↓γ⁻]]$$

\subsection{The Language of Types}

The types of \fexists are given in \cref{fig:declarative-types}.
They are stratified into two syntactic 
categories (polarities): positive and negative,  
similar to the values and computations in the \CBPV system.

\begin{figure}
  \begin{minipage}[t]{0.6\textwidth}
    \begin{supertabular}{ll p{0.8\textwidth}}
      \ottiNInline\\
    \end{supertabular}
  \end{minipage}
  \begin{minipage}[t]{0.35\textwidth}
    \begin{supertabular}{ll p{0.8\textwidth}}
      \ottiPInline\\
    \end{supertabular}
  \end{minipage}

  \caption{Declarative Types of \fexists}
  \label{fig:declarative-types}
\end{figure}

\begin{itemize}
\item [$\pm$] $[[na]]$ and $[[pa]]$
  denote negative and positive type variables, respectively.
  The variables are either introduced by the corresponding quantifiers 
  ($[[∀]]$ and $[[∃]]$, respectively) or declared in the type context;
\item [$-$] a function $[[iP → iN]]$ takes a \emph{positive} input $[[iP]]$ and
returns a \emph{negative} output $[[iN]]$; the function type itself is
\emph{negative}, so the `arrow' operator is right-associative: 
$[[iP → iQ → iN]]$ is equivalent to $[[iP → (iQ → iN)]]$;
\item [$-$] similarly to functions, a polymorphic abstraction $[[∀pas.iN]]$
  quantifies a negative type $[[iN]]$ over a list of positive type variables $[[pas]]$;
\item [$+$] dually, $[[∃nas.iP]]$ binds negative variables in a positive type $[[iP]]$;
\item [$\pm$] an upshift $[[↑iP]]$ and a downshift $[[↓iN]]$ change the polarity of a type
  from positive to negative and vice versa.
  At the level of terms, the constructors for shift types are
  $[[return v]]$ and $[[{c}]]$ (thunked computation), respectively.
\end{itemize}

\paragraph{Syntactic Conventions}
We always consider terms and types up to alpha-equivalence, and capture-avoiding
substitution. Besides, we assume the quantification is associative,
for example, $[[∀pas.∀pbs.iN]]$ and $[[∀pas,pbs.iN]]$
represent the same type.

\paragraph{Type Context and Type Well-formedness}

In the construction of \fexists, a type context (denoted as $[[Γ]]$) is
represented as a \emph{set} of positive and negative type variables and it is
used to assert the well-formedness of types. The well-formedness (or
well-scopeness) of a type is denoted as $[[Γ ⊢ iP]]$ and $[[Γ ⊢ iN]]$ and it
asserts that all type variables are either bound by a quantifier ($[[∀]]$ and
$[[∃]]$) or declared in the context $[[Γ]]$. The well-formedness checking is an
\emph{algorithmic} procedure. We represent it as a system of inference rules,
that correspond to a recursive algorithm taking the context and the type as
input. For brevity, we omit these rules, as they are standard binder scoping
rules.

\subsection{The Language of Terms}

\begin{figure}
  \begin{supertabular}{ll p{0.8\textwidth}}
    \ottcInline\\
    \ottvInline\\
  \end{supertabular}
  \caption{Declarative Terms of \fexists}
  \label{fig:declarative-terms}
\end{figure}

In \cref{fig:declarative-terms}, we define the language of terms of \fexists.
The language is a combination of \systemf and \CBPV constructs: the
terms are stratified into values and computations; there are return 
and thunk constructors, as well as several forms of let bindings to
sequentialize the computation/application with or without type annotation;

\begin{itemize}
    \item [$+$] $[[x]]$ denotes a term variable.
      Following the \CBPV stratification, we only have \emph{positive} (value)
      term variables;
    \item [$+$] $[[{c}]]$ is a value corresponding to a thunked 
        or suspended computation;
    \item [$\pm$] $[[(c : iN)]]$ and $[[(v : iP)]]$ allow one to annotate 
        positive and negative terms;
    \item [$-$] $[[return v]]$ is a pure computation, returning a value;
    \item [$-$] $[[λ x : iP . c]]$ and $[[Λ pa . c]]$
        are standard lambda abstractions. Notice that we require
        the type annotation for the argument of $[[λ]]$;
    \item [$-$] $[[ let x = v ; c]]$ is a standard let, binding
        a value $[[v]]$ to a variable $[[x]]$ in a computation $[[c]]$;
    \item [$-$] computational let form $[[ let x : iP = c; c']]$ 
      operates similarly to a monadic bind.
      It binds the result of a computation $[[c : ↑iP]]$ 
      to a variable $[[x : iP]]$, and continues with a computation $[[c']]$;
    \item [$-$] applicative let forms $[[let x : iP = v ( args ) ; c]]$ and
        $[[let x = v ( args ) ; c]]$ can be viewed as a special case of the
        computational let, when the first computation is a function application
        $[[v]] ([[args]])$.
        They take a suspended computation $[[v]]$, apply it to a list of
        arguments, bind the (pure) result to a variable
        $[[x]]$, and continues with a computation $[[c]]$. If the resulting type
        of the application is unique, one can omit the type annotation, as in
        the second form: it will be inferred by the algorithm;
    \item [$-$] $[[let∃ ( nas , x ) = v ; c]]$
        is the standard eliminator of an existential type (unpack):
        expecting $[[v]]$ to be of an existential type,
        it binds the packed negative types to a list of 
        variables $[[nas]]$, binds the body of the existential
        to $[[x]]$, and continues with a computation $[[c]]$.
\end{itemize}

\paragraph{Missing constructors}
Notice that the language does not have first-class applications: 
their role is played by the applicative let forms, binding 
the result of a \emph{fully applied} function to a variable.
Also notice that the language does not have a type application (i.e. the eliminator of $[[∀]]$) and dually, it does not have \pack (i.e. the constructor of $[[∃]]$).
This is because the instantiation of polymorphic and existential types is inferred by the algorithm. 
In \cref{sec:extensions}, we discuss the way to modify the system to introduce \emph{explicit} type applications.

\subsection{The Key Ideas of the Algorithm}

The inference algorithm for \fexists is \emph{local}: it has a limited scope of
the inference information and does not propagate it between far-apart branches
of the syntax tree. Nevertheless, many difficulties appear in the inference of
the combination of polymorphic and existential types. Next, we discuss the most
challenging of these difficulties and how they are addressed in the algorithm.

\paragraph{Subtyping} 
The inference algorithm's complexity mainly lies in the subtyping relation,
which is, in particular, responsible for polymorphic instantiation. It's this
subtyping relation that enables us to apply polymorphic terms in situations
where their specific instantiations are anticipated.

To check whether one type is a subtype of another, the algorithm introduces
\emph{algorithmic} type variables ($[[α̂±]], [[β̂±]], \dots$)--- the
`placeholders' representing the polymorphic variables whose instantiation is
postponed. This way, the problem of $[[∀α⁺.↑α⁺ ≤ ↑Int]]$ becomes the problem
of finding $[[α̂⁺]]$ such that $[[↑α̂⁺ ≤ ↑Int]]$.

\newcommand{\defiff}{%
  \mathrel{\vbox{\offinterlineskip\ialign{%
    \hfil##\hfil\cr
    $\scriptscriptstyle\text{def}$\cr
    %\noalign{\kern0ex}
    $\iff$\cr
}}}}

\paragraph{Equivalence}
The system introduces another complexity: the equivalence induced by subtyping
($[[Γ ⊢ iN ≈ iM]] \defiff [[Γ ⊢ iN ≤ iM]] \text{ and } [[Γ ⊢ iM ≤ iN]]$) is not
straightforward. This relation extends beyond just alpha-equivalent types. As
the neighboring polymorphic quantifiers can be instantiated in arbitrary order,
the equivalence relation must permit the \emph{permutations} of the quantifiers.
For example, we have $[[· ⊢ ∀α⁺.∀β⁺.iN ≈ ∀β⁺.∀α⁺.iN]]$. Similarly, the
equivalence permits the \emph{removal/addition} of the unused quantifiers: $[[·
⊢ ∀α⁺.↑Int ≈ ↑Int]]$.

To handle the complexity of non-trivial equivalence, we employ a type
normalization algorithm. This approach simplifies mutual subtyping to a more
manageable alpha-equivalence by eliminating unused quantifiers and reordering
the rest into a canonical permutation, effectively replacing subtyping-induced
equivalence on terms with alpha-equivalence on their normal forms.

\paragraph{Least Upper Bound}
Eventually, the subtyping algorithm reduces the initial subtyping problem to a set of
constraints that need to be resolved. At this moment, another problem arises:
as one variable can occur in multiple constraints, the resolution becomes non-trivial.
In particular, to resolve the conjunction of $[[α̂⁺ ≥ iP]]$ and $[[α̂⁺ ≥ iQ]]$,
we would need to find the least upper bound of $[[iP]]$ and $[[iQ]]$.

In many type systems, the least upper bound boils down to the trivial syntactic equality.
Because of that, the set of constraints is typically resolved using \emph{unification}.
However, this problem is trickier in the presence of existential types. For example, 
in \fexists, the least upper bound of $[[↓↑Int]]$ and $[[↓↑Bool]]$ is $[[∃α⁻.↓α⁻]]$,
because the quantified $[[α⁻]]$ can be instantiated to either $[[↑Int]]$ or $[[↑Bool]]$.

To find this least upper bound, we use \emph{anti-unification}---the dual of
unification. While unification finds \emph{the most general instance} of two
given patterns, the anti-unification finds \emph{the most detailed} (under the
restrictions of the system) \emph{pattern} that matches two given types. In the
example above, the anti-unifier of $[[↓↑Int]]$ and $[[↓↑Bool]]$ is $[[↓α̂⁻]]$,
and in \fexists, this unifier is the most specific (in particular because
$[[↓↑α̂⁺]]$ is not considered as only \emph{negative} variables are
existentially quantified).

\paragraph{Impredicativity}
Another difficulty that hinders the constraint resolution arises from the
impredicativity of the system---the ability to quantify over the types that are 
themselves polymorphic. In the impredicative polymorphic system (even
with only $[[∀]]$-quantifiers), the naturally formulated subtyping is
undecidable, which follows from the undecidability of second-order unification.
All the more the undecidability is evident in the presence of existential types
and subtyping constraints. For instance, in \fexists the constraint 
$[[α̂⁺ :≥ ↓↑α̂⁺]]$ would have a surprising impredicative solution $[[α̂⁺]] = [[∃α⁻.↓α⁻]]$.

To maintain the decidability in the presence of impredicativity, we carefully
restrict the system using the polarity stratification. One of the main
constraints that we put on subtyping is the invariance of the shift operators:
the subtyping $[[↑iP ≤ ↑iQ]]$ is only allowed if $[[iQ ≥ iP]]$
\emph{and} $[[iP ≥ iQ]]$. These constraints do not trivialize the system: we
still have non-trivial upper bounds, and thus, need to use anti-unification to
find them. However, it prevents such examples as $[[α̂⁺ :≥ ↓↑α̂⁺]]$, and 
as we prove, makes the constraint resolution decidable.