### 2024-05-07 oopsla24/tex/07-conclusion.tex
edit (misc)

We have presented a type inference algorithm for an impredicative polymorphic
lambda calculus with existential types and subtyping. The system is designed in
\CBPV paradigm, which allowed us to restrict it to a decidable fragment of the
language described declaratively. We presented the inference algorithm
and proved its soundness and completeness with respect to the specification. The
algorithm combines unification---a standard type inference technique---with
anti-unification---which has not been used in type inference before.