### 2024-06-11 paper/tex/03-declarative-system.tex

\label{sec:declarative-system}

In this section, we present the declarative system of \fexists,
which serves as a specification of the type inference algorithm.
The declarative system is represented as a set of inference rules
and consists of two main subsystems: subtyping and type inference.
First, we present the declarative subtyping rules specifying 
when one type is a subtype of another; next, we discuss 
the equivalence relation induced by mutual subtyping;
finally, we present the type inference rules,
that refer to the subtyping and equivalence relation.
We conclude this section by discussing the relation between the
proposed type system and the standard \systemf.

\subsection{Subtyping}

\begin{figure}
  \begin{minipage}[t]{0.49\textwidth}
    \ottdefnDOneNsubLabeled{}
  \end{minipage}%
  \begin{minipage}[t]{0.49\textwidth}
    \ottdefnDOnePsupLabeled{}
  \end{minipage}
  \hfill\\
  \begin{minipage}[t]{0.49\textwidth}
    \ottdefnDOneNeqLabeled{}
  \end{minipage}%
  \begin{minipage}[t]{0.49\textwidth}
    \ottdefnDOnePeqLabeled{}
  \end{minipage}
  \caption{Declarative Subtyping}
  \label{fig:declarative-subtyping}
\end{figure}

The inference rules representing declarative subtyping are shown in
\cref{fig:declarative-subtyping}. Let us discuss them in more detail.

\paragraph*{Quantifiers}  
Symmetric rules \ruleref{\ottdruleDOneForallLabel} and 
\ruleref{\ottdruleDOneExistsLabel} specify 
subtyping between top-level quantified types.
Usually, the polymorphic subtyping is represented by two rules
introducing quantifiers to the left and to the right-hand side of subtyping.
For conciseness of representation, we compose these rules into one.
First, our rule extends context $[[Γ]]$ with the quantified variables 
from the right-hand side ($[[pbs]]$ or $[[nbs]]$), 
as these variables must remain abstract.
Second, it verifies that the left-hand side quantifiers
($[[pas]]$ or $[[nas]]$) can be instantiated to continue subtyping recursively. 
This is the point of non-determinism as the instantiation can be chosen arbitrarily

The instantiation of quantifiers is modeled by substitution $[[σ]]$.
The notation $[[Γ2 ⊢ σ : Γ1]]$ specifies its domain and range.
For instance, $[[Γ, pbs ⊢ σ : {pas}]]$ means that 
$[[σ]]$ maps the variables from $[[pas]]$ to (positive) types
well-formed in $[[Γ, pbs]]$.
This way, application $[[ [σ]iN ]]$ instantiates (replaces) every
$[[αi⁻]]$ in $[[iN]]$ with $[[σ]]([[αi⁻]])$.

\paragraph*{Invariant Shifts}
As mentioned above, an important restriction that we put on the subtyping system
is that subtyping of shifted types requires their equivalence, as shown in
\ruleref{\ottdruleDOneShiftDLabel} and \ruleref{\ottdruleDOneShiftULabel}. The
reason for this is that if both of these rules were relaxed to the covariant
form, the subtyping relation would become equivalent to the standard subtyping
of \systemf, which is undecidable \cite{tiuryn-urzczyn-96}. However, it is
suggested after certain changes \ruleref{\ottdruleDOneShiftULabel} can be
relaxed to the covariant form, thereby increasing the expressiveness of the
system. These changes are discussed in \cref{sec:weakening-invariant}.

\paragraph*{Functions}
Standardly, subtyping of function types is covariant in the return type
and contravariant in the argument type.

\paragraph*{Variables}
Subtyping of variables is defined reflexively, which is enough to ensure the
reflexivity of subtyping in general. The algorithm---specifically the least
upper bound procedure---will use the fact that the subtypes of a variable
coincide with its supertypes (\cref{prop:var-no-subtypes}),
which however does not hold for arbitrary types.

\subsubsection{Properties of Declarative Subtyping}
\label{sec:decl-subtyping-properties}

A property that is important for the subtyping algorithm, 
in particular for the type \emph{upgrade} procedure (\cref{sec:lub}),
is the preservation of free variables by subtyping.
Informally, it says that the free variables
of a positive type cannot disappear in its subtypes,
and the free variables of a negative type
cannot disappear in its supertypes.

\begin{property}[Subtyping preserves free variables]
  \label{prop:subtyping-preserves-fv}
  Let us assume that all the mentioned types are well-formed in $[[Γ]]$. Then
  $[[Γ ⊢ iN1 ≤ iN2]]$ implies $[[fv(iN1) ⊆ fv(iN2)]]$,
  and $[[Γ ⊢ iP1 ≥ iP2]]$ implies $[[fv(iP1) ⊆ fv(iP2)]]$.
\end{property}

\begin{property}[Variable subtyping is trivial]
  \label{prop:var-no-subtypes}
  A subtype or a supertype of a variable is equivalent to the variable itself:

  \begin{tabular}{@{}llclcl@{}}
    $-$ & $[[Γ ⊢ α⁻ ≤ iN]]$ & $\iff$ & $[[Γ ⊢ iN ≤ α⁻]]$ & $\iff$ & $[[iN]] = [[∀pbs.α⁻]]$ \\
    $+$ & $[[Γ ⊢ α⁺ ≥ iP]]$ & $\iff$ & $[[Γ ⊢ iP ≥ α⁺]]$ & $\iff$ & $[[iP]] = [[∃nbs.α⁺]]$
  \end{tabular}
\end{property}

Another property that we extensively use is that subtyping is reflexive and transitive,
and agrees with substitution.

\begin{property}[Subtyping forms a preorder]
  For a fixed context $[[Γ]]$, the negative subtyping relation 
  $[[Γ ⊢ iN1 ≤ iN2]]$ and the positive subtyping relation 
  $[[Γ ⊢ iP1 ≥ iP2]]$
  are reflexive and transitive on types well-formed in $[[Γ]]$.
\end{property}

\begin{property}[Subtyping agrees with substitution]
  Suppose that  $[[σ]]$ is a substitution such that $[[Γ2 ⊢ σ : Γ1]]$. 
  Then
  \begin{enumerate}
    \item [$-$] $[[Γ1 ⊢ iN ≤ iM]]$ implies $[[Γ2 ⊢ [σ]iN ≤ [σ]iM]]$, and
    \item [$+$] $[[Γ1 ⊢ iP ≥ iQ]]$ implies $[[Γ2 ⊢ [σ]iP ≥ [σ]iQ]]$.
  \end{enumerate}
\end{property}

Moreover, any two \emph{positive} types have a least upper bound, which makes
positive subtyping a semilattice. The positive least upper bound can be found
algorithmically, which we will discuss in the next section.

\begin{property}[Positive least upper bound exists]
  Suppose that $[[iP1]]$ and $[[iP2]]$ are positive types
  well-formed in $[[Γ]]$.
  Then there exists a least common supertype--- a type $[[iP]]$ such that
  \begin{itemize}
    \item $[[Γ ⊢ iP ≥ iP1]]$ and $[[Γ ⊢ iP ≥ iP2]]$, and 
    \item for any $[[iQ]]$ such that $[[Γ ⊢ iQ ≥ iP1]]$ and $[[Γ ⊢ iQ ≥ iP2]]$,
      $[[Γ ⊢ iQ ≥ iP]]$.
  \end{itemize}
\end{property}

\paragraph*{Negative greatest lower bound might not exist}
The symmetric construction---the greatest lower bound of two negative types---does 
not always exist, as the following counterexample shows.
Consider the following types: 
\begin{itemize}
  \item $[[iN]]$ and $[[iQ]]$ are arbitrary closed types, 
  \item $[[iP]]$, $[[iP1]]$, and $[[iP2]]$ are non-equivalent closed types 
    such that $[[· ⊢ iP1 ≥ iP]]$ and $[[· ⊢ iP2 ≥ iP]]$, and 
    none of the types is equivalent to $[[iQ]]$.
\end{itemize}
What is the greatest common subtype of  
$[[iQ → ↓↑iQ → ↓↑iQ → iN]]$ and $[[iP → ↓↑iP1 → ↓↑iP2 → iN]]$?
The type $[[∀α⁺,β⁺,γ⁺. α⁺ → ↓↑β⁺ → ↓↑γ⁺ → iN]]$ is a common subtype,
however, it is not the greatest one, as it is too general.

One can find two greater candidates:
$[[iM1 = ∀α⁺,β⁺. α⁺ → ↓↑α⁺ → ↓↑β⁺ → iN]]$ and $[[iM2 = ∀α⁺,β⁺. β⁺ → ↓↑α⁺ → ↓↑β⁺ → iN]]$.
Instantiating $[[α⁺]]$ and $[[β⁺]]$ with $[[iQ]]$ ensures 
that both of these types are subtypes of $[[iQ → ↓↑iQ → ↓↑iQ → iN]]$;
instantiating $[[α⁺]]$ with $[[iP1]]$ and $[[β⁺]]$ with $[[iP2]]$
demonstrates the subtyping with $[[iP → ↓↑iP1 → ↓↑iP2 → iN]]$,
as $[[iP]]$ is a subtype of both $[[iP1]]$ and $[[iP2]]$.

By analyzing the inference rules, we can prove that
both $[[iM1]]$ and $[[iM2]]$ are \emph{maximal} common 
subtypes, \ie there is no common subtype that is greater than them.
However, $[[iM1]]$ and $[[iM2]]$ are not equivalent,
which means that none of them is the greatest.

\subsection{Equivalence and Normalization}
\label{sec:decl-equivalence}

The subtyping relation forms a preorder on types,
and thus, it induces an equivalence relation \aka bicoercibility 
\cite{tiuryn1995:bicoercibility}.
The declarative specification of subtyping must be defined up to this equivalence.
Moreover, the algorithms we use must withstand changes in input types within the equivalence class.
To deal with non-trivial equivalence, 
we use normalization---a function that uniformly selects a representative of the equivalence class.

Using normalization gives us two benefits:
\begin{enumerate*}
  \item [(i)] we do not need to modify significantly standard operations such as unification to withstand non-trivial equivalence, and
  \item [(ii)] if the subtyping (and thus, the equivalence) changes, we only need to modify the normalization function, 
    while the rest of the algorithm remains the same.
\end{enumerate*}

In our system, equivalence is richer than equality. Specifically,
while staying within one equivalence class, one can 
change the type:
\begin{enumerate}
  \item[(i)] introduce and remove redundant quantifiers. For example, $[[∀α⁺,β⁺.↑α⁺]]$ is equivalent
  but not equal to $[[∀α⁺.↑α⁺]]$;
  \item[(ii)] reorder adjacent quantifiers. For example, $[[∀α⁺,β⁺.α⁺ → β⁺ → γ⁻]]$ is equivalent but not equal to
  $[[∀α⁺,β⁺.β⁺ → α⁺ → γ⁻]]$; 
  \item[(iii)] make the transformations (i) and (ii) at any position in the type.
\end{enumerate}

It turns out that the transformations (i-iii) are complete, 
in the sense that they generate the whole equivalence class.
This way, to normalize the type, one must
\begin{enumerate}
  \item [(i)] remove the redundant quantifiers,
  \item [(ii)] reorder the quantifiers to the canonical order,
  \item [(iii)] do the procedures (i) and (ii) recursively on the subterms.
\end{enumerate}

The normalization algorithm is shown in \cref{fig:type-nf}. The steps (i-ii) are
implemented by the \emph{ordering} function `$[[ord varset in iN]]$' and `$[[ord
varset in iP]]$'. For a set of variables $[[varset]]$, and a type, it returns a
list of variables from $[[varset]]$ that occur in the type in the \emph{order of
their first occurrence}. For brevity, we omit the formal definition of ordering
referring the reader to the appendix (\cref{appendix}).

\begin{figure}[h]

  \begin{multicols}{2}
  \ottdefnNrmNNormLabeled{}
  \\
  $[[ord varset in iN]]$ returns a list of variables from 
  $[[varset ∩ fv(iN)]]$ in the order of their first occurrence in $[[iN]]$
  \columnbreak\\
  \ottdefnNrmPNormLabeled{}
  \\
  $[[ord varset in iP]]$ returns a list of variables from 
  $[[varset ∩ fv(iP)]]$ in the order of their first occurrence in $[[iP]]$
  \end{multicols}

  \caption{Type Normalization Procedure} 
  \label{fig:type-nf}
\end{figure}

For the normalization procedure, 
we prove soundness and completeness \wrt the equivalence relation.
\begin{property}[Correctness of normalization]
  Assuming all types are well-formed in $[[Γ]]$,
  \begin{itemize}
    \item[$-$] $[[Γ ⊢ iN ≈ iM]]$ is equivalent to $[[nf(iN) = nf(iM)]]$, and
    \item[$+$] $[[Γ ⊢ iP ≈ iQ]]$ is equivalent to $[[nf(iP) = nf(iQ)]]$.
  \end{itemize}
\end{property}

\subsection{Typing}

The declarative specification of the type inference is given in 
\cref{fig:declarative-inference}.
The positive typing judgment $[[Γ ; Φ ⊢ v : iP]]$ is read as 
``under the type context $[[Γ]]$ and variable context $[[Φ]]$,
the term $[[v]]$ is allowed to infer type $[[iP]]$'',
where $[[Φ]]$---the variable context---is defined standardly as
a set of pairs of the form $[[x : iP]]$. 
The negative typing judgment is read similarly.

The \emph{Application typing} judgment
infers the type of the application of a function to a list of arguments.
It has form $[[Γ ; Φ ⊢ iN ● args ⇒> iM]]$, 
which reads ``under the type context $[[Γ]]$ and variable context $[[Φ]]$,
the application of a function of type $[[iN]]$ to the list of arguments $[[args]]$
is allowed to infer type $[[iM]]$''.

\begin{figure}[t!]
  \ottdefnsDTLabeled
  \caption{Declarative Inference}
  \label{fig:declarative-inference}
\end{figure}

Let us discuss the rules of the declarative system in more detail.

\paragraph*{Variables}
  Rule \ruleref{\ottdruleDTVarLabel} allows us to infer
  the type of a variable from the context. 
  In literature, one can find another version of this rule,
  that enables inferring a type \emph{equivalent}
  to the type from the context. 
  In our case, the inference of equivalent types
  is admissible in general case by \ruleref{\ottdruleDTPEquivLabel}.

\paragraph*{Annotations}
  The annotation rules \ruleref{\ottdruleDTNAnnotLabel}
  and \ruleref{\ottdruleDTPAnnotLabel} use subtyping.
  The annotation is only valid if the
  inferred type is a subtype of the annotation type.

\paragraph*{Abstractions}
  The typing of lambda abstraction is standard. 
  Rule \ruleref{\ottdruleDTtLamLabel} first checks
  that the given type annotating the argument is well-formed,
  and then infers the type of the body in the extended context.
  As a result, it returns an arrow type of function from the
  annotated type of the argument to the type of the body.
  Rule \ruleref{\ottdruleDTTLamLabel} infers polymorphic $[[∀]]$-type. 
  It extends the type context with the quantifying variable $[[pa]]$ and 
  infers the type of the body. As a result, it returns a polymorphic type
  quantifying the abstracted variable $[[pa]]$ over the type of the body.

\paragraph*{Return and thunk}
  Rules \ruleref{\ottdruleDTReturnLabel} 
  and \ruleref{\ottdruleDTThunkLabel}
  simply add the corresponding shift constructors to the type of the body.

\paragraph*{Unpack}
  Rule \ruleref{\ottdruleDTUnpackLabel} types elimination of $[[∃]]$.
  First, it infers the normalized type of the existential package.
  The normalization is required to fix the order of the quantifying variables
  to bind them. Alternative approaches that do not require normalization
  will be discussed in \cref{sec:explicit-type-application}.
  After the bind, the rule infers the type of the body 
  and checks that it does not use the bound variables so that they do not 
  escape the scope.

\paragraph*{Let binders}
  Rule \ruleref{\ottdruleDTVarLetLabel} is a standard rule for typing let
  binders: we infer the type of the bound value and continue the typing of the
  computation in the extended context.

  Rule \ruleref{\ottdruleDTCVarLetLabel} associates a variable with a computation.
  The inferred type for this computation ($[[iM]]$) must be castable to a
  shifted positive type $[[↑iP]]$, with $[[iP]]$ then assigned to the bound
  variable $[[x]]$ to type the let binder's body. Like all annotated
  constructors, we also verify the annotation type $[[iP]]$'s well-formedness.

\paragraph*{Applicative let binders}
  Rules \ruleref{\ottdruleDTAppLetLabel} and \ruleref{\ottdruleDTAppLetAnnLabel}
  infer the type of the applicative let binders.
  Both of them infer the type of the head $[[v]]$ 
  and invoke the application typing to infer the type of the application 
  before recursing on the body of the let binder.

  The former rule infers the type of an \emph{unannotated} let binder, and thus
  it requires the resulting type of application to be \emph{the principal type},
  so that the type we assign to the bound variable $[[x]]$ is determined.
  In this context, principality means minimality. In other words, 
  $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ principal]]$ means that
  any other type $[[iQ']]$ inferrable for the application (\ie $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ']]$)
  is greater than the principal type $[[iQ]]$, \ie $[[Γ ⊢ iQ' ≥ iQ]]$.

  The latter rule is for the \emph{annotated} binder,
  and thus, the type of the bound $[[x]]$ is given, 
  however, the rule must check that this type is a
  a supertype of the inferred type of the application. 
  This check is done by invoking the subtyping judgment
  $[[Γ ⊢ iM' ≤ ↑iP]]$.
  % This judgment is more restrictive than checking bare 
  % $[[Γ ⊢ iP ≥ iQ]]$, however, it is necessary
  % to make the algorithm complete as it allows us to preserve
  % certain invariants throughout the algorithm to ensure
  % the decidability of the produced unification task.
  % In \cref{sec:proof-subtyping}, we discuss these invariants in
  % more detail, and in \cref{sec:weakening-invariant} we suggest 
  % a way to relax them.

\paragraph*{Typing up to equivalence}
  As discussed in \cref{sec:decl-equivalence}, the subtyping, as a preorder, 
  induces a non-trivial equivalence relation on types. 
  The system must not distinguish between equivalent types,
  and thus, type inference must be defined up to equivalence. 
  For this purpose, we use rules \ruleref{\ottdruleDTPEquivLabel}  
  and \ruleref{\ottdruleDTNEquivLabel}.
  They allow one to replace the inferred type with an equivalent one.

\paragraph*{Application to an empty list of arguments}
  The base case of the application type inference is 
  represented by rule \ruleref{\ottdruleDTEmptyAppLabel}.
  If the head of the type $[[iN]]$ is applied to no arguments, 
  the type of the result is allowed to be $[[iN]]$ or any 
  equivalent type. We need to relax this rule up to equivalence
  to ensure the corresponding property globally:
  the inferred application type can be replaced with an equivalent one.
  Alternatively, we could have added a separate rule similar to 
  \ruleref{\ottdruleDTPEquivLabel}, however, 
  the local relaxation is sufficient to prove the global property.

\paragraph*{Application of a polymorphic type $[[∀]]$}
  The complexity of the system lives in the rules whose output type is not
  immediately defined by their input and the output of their premises (\aka not
  mode-correct \cite{dunfield2020:bidirectional}). In our typing system,
  \ruleref{\ottdruleDTForallAppLabel} is such a rule: the instantiation of the
  quantifying variables is not known a priori. The algorithm we present in
  \cref{sec:algorithm} delays this instantiation until more information about it
  (in particular, typing constraints) is collected.

  To ensure the priority of application between this rule and 
  \ruleref{\ottdruleDTEmptyAppLabel}, we also check that 
  the list of arguments is not empty.

\paragraph*{Application of an arrow type}
  Another application rule is \ruleref{\ottdruleDTArrowAppLabel}.
  This is where the subtyping is used to check that the type of the argument
  is convertible to (a subtype of) the type of the function parameter.
  In the algorithm (\cref{sec:algorithm}), this subtyping check will provide the constraints
  we need to resolve the delayed instantiations of the quantifying variables.

  \vspace{\baselineskip}
% \subsubsection{Properties of Declarative Typing}
  An important property that the declarative system has is
  that the declarative specification is correctly defined for
  equivalence classes.

\begin{property}[Declarative typing is defined up to equivalence]
  Let us assume that $[[Γ ⊢ Φ1 ≈ Φ2]]$, \ie  
  the corresponding types assigned by $[[Φ1]]$ and
  $[[Φ2]]$ are equivalent in $[[Γ]]$.
  Also, let us assume that 
  $[[Γ ⊢ iN1 ≈ iN2]]$, $[[Γ ⊢ iP1 ≈ iP2]]$,
  and $[[Γ ⊢ iM1 ≈ iM2]]$. Then
  \begin{itemize}
    \item [$-$] $[[Γ ; Φ1 ⊢ c : iN1]]$ holds if and only if $[[Γ ; Φ2 ⊢ c : iN2]]$,
    \item [$+$] $[[Γ ; Φ1 ⊢ v : iP1]]$ holds if and only if $[[Γ ; Φ2 ⊢ v : iP2]]$, and
    \item [$\bullet$] $[[Γ; Φ1 ⊢ iN1 ● args ⇒> iM1]]$ holds if and only if $[[Γ; Φ2 ⊢ iN2 ● args ⇒> iM2]]$.
  \end{itemize}
\end{property}

\subsection{Relation to \systemf}
\label{sec:rel-to-systemf}

To establish a correspondence between \fexists and standard unpolarized \systemf,
we present a translation in both ways: the polarization from \systemf to \fexists
and the depolarization from \fexists to \systemf.

\paragraph*{Type-level Translation}

The type depolarization (\cref{fig:depolarization}) simply forgets the polarization structure of types:
it removes the shift operators and the polarities of the free type variables.

The type polarization (\cref{fig:polarization}) is more involved, as there are
multiple ways to polarize a type: for instance, a variable $[[α]]$ can be
polarized to $[[α⁺]]$ or $[[α⁻]]$. The choice of polarization affects the
execution strategy of the program. 
In particular, a functional type can be represented in either positive (thunked, call-by-name) way:
$[[↓(iP → ↑iQ)]]$ or negative (call-by-value): $[[↓iN → iM]]$.
We chose the positive polarization: every \systemf type is translated to a
\emph{positive} type of \fexists. This choice makes it smoother to lift the
translation to the term level, which will be discussed next.

The type-level translation is naturally lifted to the contexts:
$[[|Γ|]]$ is the context $[[Γ]]$ with every type depolarized,
and symmetrically, $[[+|fΓ|+]]$ is the context $[[fΓ]]$ with every type polarized.
The context translation is used to formulate the soundness of the
 term-level translation, which we discuss next.

\begin{figure}[ht]
  \begin{minipage}[t]{0.65\linewidth}
    \centering
      \begin{minipage}[t]{0.32\linewidth}
      \ottfundefnunpolP{}
      \end{minipage}
      \begin{minipage}[t]{0.32\linewidth}
      \ottfundefnunpolN{}
      \end{minipage}
    \caption{Type Depolarization \protect\footnotemark}
    \label{fig:depolarization}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[t]{0.3\linewidth}
    \centering
    \ottfundefnpolarP{}
    \caption{Type Polarization}
    \label{fig:polarization}
  \end{minipage}

\end{figure}

\footnotetext{Here the \systemf existential $[[∃ as . fT]]$ is a syntax sugar for its standard encoding $[[∀b.(∀as.(fT → b)) → b]]$}

\paragraph*{Term-level Translation}

The term-level translations between the systems are more complex. They are
defined as \emph{elaborations} of the typing derivations. In other words, the
input of the translation is not a term, but a whole typing derivation in the
corresponding system. For example, the subtyping elaboration $[[Γ ⊢ iN ≤ iM ⤳
ft]]$ constructs a \systemf function $[[ft]]$ of type $[[|iN| → |iM|]]$ that
witnesses the given \fexists-subtyping $[[Γ ⊢ iN ≤ iM]]$; a typing elaboration
$[[Γ; Φ ⊢ v : iP ⤳ ft]]$ builds a term $[[ft]]$---a \systemf counterpart of
$[[v]]$---of type $[[|iP|]]$.

For brevity, we omit the formal definition of the term-level elaboration,
referring the reader to the appendix (\cref{sec:term-translation}). We define
the elaboration in such a way that it preserves the soundness invariants given
in \cref{tab:elaboration-soundness}.

\begin{table}[h]
\centering
\begin{tabular}{rll}
\hline
\multicolumn{2}{c}{\textbf{Elaboration Judgment}} & \textbf{The Soundness Property} \\
\hline
Negative Subtyping & $[[Γ ⊢ iN ≤ iM ⤳ ft]]$ & $[[|Γ| ; · ⊢ ft : |iN| → |iM| ]]$ \\
Positive Subtyping & $[[Γ ⊢ iP ≥ iQ ⤳ ft]]$ & $[[|Γ| ; · ⊢ ft : |iQ| → |iP| ]]$ \\
Negative Typing & $[[Γ; Φ ⊢ c : iN ⤳ ft]]$ & $[[|Γ| ; |Φ| ⊢ ft : |iN|]]$ \\
Positive Typing & $[[Γ; Φ ⊢ v : iP ⤳ ft]]$ & $[[|Γ| ; |Φ| ⊢ ft : |iP|]]$ \\
Application Typing & $[[Γ ; Φ ⊢ iN ● args ⇒> iM ⤳ fe; fargs]]$ & $[[|Γ| ; |Φ|, x:|iN| ⊢ fe (x ● ft *) : |iM|]]$ \\
\hline
\systemf Typing & $[[fΓ ; fΦ ⊢ ft : fT ⤳ c]]$ & $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|fT|+]]$ \\
\hline
\end{tabular}
\caption{Soundness of the Elaboration:
  the elaboration judgment on the left implies the typing judgment on the right}
  \label{tab:elaboration-soundness}
\end{table}

The least obvious aspect of the elaboration soundness is the application typing.
The elaboration judgment $[[Γ ; Φ ⊢ iN ● args ⇒> iM ⤳ fe; fargs]]$ 
outputs two things: 
\begin{enumerate*}
  \item [(i)] $[[fargs]]$---the \systemf counterpart of the argument list $[[args]]$, and 
  \item [(ii)] $[[fe]]$---a \systemf term that is used for the final cast 
    of the application $[[(x ● ft *)]]$ to the resulting type $[[|iM|]]$.
\end{enumerate*}
This final cast is needed to 
reflect the \fexists feature of \ruleref{\ottdruleDTEmptyAppLabel}
that permits the equivalent conversion of the resulting type $[[Γ ⊢ iN ≈ iN']]$,
as \fexists-equivalent types
might have different \systemf representations (\ie $[[|iN|]] \neq [[|iN'|]]$).

% \begin{figure}[h]
%   \begin{equation*}
%   \begin{aligned}
%     [[|↓iN|]] &= [[|iN|]] \\
%     [[|↑iP|]] &= [[|iP|]] \\
%   \end{aligned}
%   \qquad
%   \begin{aligned}
%     [[|α±|]] &= [[a]] \\
%     [[|iP → iN|]] &= [[|iP| → |iN|]] \\
%   \end{aligned}
%   \qquad
%   \begin{aligned}
%     [[|∀α⁺.iN|]] &= [[∀α.|iN|]] \\
%     [[|∃α⁻.iP|]] &= [[∃α.|iP|]] \\
%   \end{aligned}
%   \end{equation*}
%   \caption{Type depolarization}
%   \label{fig:depolarization}
% \end{figure}

% The translation from \systemf to \fexists can be done in two canonical ways
% depending on the desired polarity. The polarity will also determine the 
% execution strategy \ilyam{how?}.
% Moreover, the translation procedure depends on the \emph{polarity} of the
% free type variables. Thus, the translation functions 
% $[[-| fT |-^Γ]]$ and $[[+| fT |+^Γ]]$ are indexed by a context $[[Γ]]$, 
% containing either $[[α⁺]]$ or $[[α⁻]]$ for each variable $[[α]]$ free in $[[fT]]$.

% \begin{figure}[h]
%   \begin{equation*}
%   \begin{aligned}
%     [[-| a |-^Γ]] &= [[α⁻]]  \text{,~if $[[α⁻ ∊ Γ]]$}\\
%     [[-| a |-^Γ]] &= [[↑α⁺]] \text{,~if $[[α⁺ ∊ Γ]]$}\\
%     [[-| fA → fB |-^Γ]] &= [[ +|fA|+^Γ → -|fB|-^Γ  ]] \\
%     [[-| ∀α.fT |-^Γ]] &= [[ ∀α⁺.-|fT|-^Γ,α⁺ ]] \\
%     [[-| ∃α.fT |-^Γ]] &= [[ ↑ ∃α⁻.+|fT|+^Γ,α⁻ ]] \\
%   \end{aligned}
%   \qquad
%   \begin{aligned}
%     [[+| a |+^Γ]] &= [[α⁺]]  \text{,~if $[[α⁺ ∊ Γ]]$}\\
%     [[+| a |+^Γ]] &= [[↓α⁻]] \text{,~if $[[α⁻ ∊ Γ]]$}\\
%     [[+| fA → fB |+^Γ]] &= [[ ↓ +|fA|+^Γ → -|fB|-^Γ  ]] \\
%     [[+| ∀α.fT |+^Γ]] &= [[ ↓∀α⁺.-|fT|-^Γ,α⁺ ]] \\
%     [[+| ∃α.fT |+^Γ]] &= [[ ∃α⁻.+|fT|+^Γ,α⁻ ]] \\
%   \end{aligned}
%   \end{equation*}
%   \caption{Type polarization}
%   \label{fig:polarization}
% \end{figure}